---
title: "Brucellosis Codes"
format:
  html:
    toc: true
    toc-location: left
    code-fold: true
    code-summary: "Show the code"
editor: visual
---

## Data Processing

```{r}
source('clean_all_data.R')
```

```{r Grouping the Data}

# Identifying the outliers in the number of cases
numeric_columns <- df_incidence2.1 %>%
  dplyr::select(contains('cases')) |> 
  keep(is.numeric)

numeric_columns %>%
  map(~summary(.))

numeric_columns %>%
  imap(
    ~ ggplot(data = data.frame(y = .x)) +
      geom_boxplot(aes(x = 1, y = y)) +
      labs(title = .y) +
      theme_light()
  )

# Replacing the outliers with the mean
df_incidence2 <- df_incidence2.1 |> 
  mutate(
    catt_cases = ifelse(catt_cases >= 69, round(mean(catt_cases, na.rm = T)), catt_cases),
    goat_cases = ifelse(goat_cases > 28, round(mean(goat_cases, na.rm = T)), goat_cases)
  ) |> 
  mutate(date = as.Date(date))

df_tot_cases <- df_incidence2 |>
  group_by(date, county) |>
  summarise(across(contains("cases"), ~ sum(., na.rm = T))) 

# Population per year, per county
df_pop <- df_incidence2 |> 
  dplyr::select(date, county, contains("pop")) %>%
  distinct(.) |>
  as_tibble() |>
  group_by(date, county) %>%
  summarise(across(where(is.numeric), ~unique(.)))

df_1 <- df_tot_cases |>
  merge(df_pop, by = c("date", "county")) |> 
  filter(!is.na(date)) |>
  mutate(
    human_incidence = round((hum_cases / pop) * 1000, 4),
    catt_incidence = round((catt_cases / catt_pop) * 1000000, 4),
    cam_incidence = round((cam_cases / cam_pop) * 1000000, 4),
    goat_incidence = round((goat_cases / goat_pop) * 1000000, 4),
    shp_incidence = round((shp_cases / sheep_pop) * 1000000, 4)
  ) |>
  dplyr::select(date, county, contains(c("incidence", "cases"))) |>
  as_tibble() |> 
  arrange(county)

df_1_pop <- df_tot_cases |>
  merge(df_pop, by = c("date", "county")) |> 
  dplyr::select(date, county, pop)
write.csv(df_1_pop, 'county_humanpop.csv', row.names = F)

df_1_complete <- df_1 |>
  dplyr::select(
    date,
    county,
    human_incidence,
    catt_incidence,
    goat_incidence,
    hum_cases,
    catt_cases,
    goat_cases
  ) |>
  filter(catt_incidence > 0)
write.csv(df_1_complete, "df_1_complete.csv", row.names = F)

# Outliers
numeric_columns <- df_1 %>%
  dplyr::select(contains('incidence')) |> 
  keep(is.numeric)

numeric_columns %>%
  map(~summary(.))

numeric_columns %>%
  imap(
    ~ ggplot(data = data.frame(y = .x)) +
      geom_boxplot(aes(x = 1, y = y)) +
      labs(title = .y) +
      theme_light()
  )


df_cum <- df_tot_cases |>
  merge(df_pop, by = c("date", "county")) |> 
  filter(!is.na(date)) |> 
  rowwise() |> 
  mutate(
    animal_cases = sum(catt_cases, goat_cases, shp_cases, cam_cases, na.rm = T),
    animal_pop = sum(catt_pop, goat_pop, sheep_pop, cam_pop, na.rm = T),
    animal_cases = ifelse(animal_cases == 0, NA, animal_cases),
    animal_incidence = round((animal_cases / animal_pop) * 1000000, 4),
    human_incidence = round((hum_cases / pop) * 1000, 4)
  ) |> 
  dplyr::select(date, county, contains(c("incidence", 'cases'))) |> 
  as_tibble() |> 
  mutate()

df_cum_complete <- df_cum |> 
  dplyr::select(date, county, human_incidence, animal_incidence, animal_cases, hum_cases) |> 
  filter(animal_incidence > 0)
write.csv(df_cum_complete, "df_cum_complete.csv", row.names = F)

## County
df_county <- df_incidence2.1 |> 
  dplyr::select(county, contains("cases")) |> 
  mutate(
    catt_cases = ifelse(catt_cases >= 69, round(mean(catt_cases, na.rm = T)), catt_cases),
    goat_cases = ifelse(goat_cases > 28, round(mean(goat_cases, na.rm = T)), goat_cases)
  ) |> 
  group_by(county) |> 
  summarise(across(where(is.numeric), ~round(sum(., na.rm = T))))
```

```{r plotting the trends over time}

df_incidence2_trend <- df_incidence2.1 |> 
  dplyr::select(date, 
         county, 
         diseases, 
         diagnosis, 
         contains("cases"), 
         catt_pop, 
         goat_pop, 
         sheep_pop, 
         cam_pop) |>
  mutate(year = year(as.Date(date))) |> 
  merge(pop, by = c("county", "year")) 

df_tot_cases_trend <- df_incidence2_trend |>
  group_by(date) |>
  summarise(across(contains("cases"), ~ sum(., na.rm = T))) |> 
  mutate(across(contains('cases'), ~ifelse(. == 0, NA, .)))

df_tot_pop_trend <- df_incidence2_trend |>
  dplyr::select(date, county, contains("pop")) %>%
  distinct(.) |>
  as_tibble() |>
  group_by(date) %>%
  summarise(
    sheep_pop = sum(sheep_pop),
    goat_pop = sum(goat_pop),
    cam_pop = sum(cam_pop),
    catt_pop = sum(catt_pop),
    hum_pop = sum(pop)
  ) %>%
  as_tibble()


df_1_trend <- df_tot_cases_trend |>
  merge(df_tot_pop_trend, by = "date") |>
  filter(!is.na(date)) |>
  mutate(
    human_incidence = round((hum_cases / hum_pop) * 1000, 4),
    catt_incidence = round((catt_cases / catt_pop) * 1000000, 4),
    cam_incidence = round((cam_cases / cam_pop) * 1000000, 4),
    goat_incidence = round((goat_cases / goat_pop) * 1000000, 4),
    shp_incidence = round((shp_cases / sheep_pop) * 1000000, 4),
    date = as.Date(date)
  ) |>
  dplyr::select(date, contains("incidence")) |>
  as_tibble() |> 
  mutate(
    
  )

write.csv(df_1_trend, "individual_incidence.csv")

df_1_trend_complete <- df_1_trend |>
  dplyr::select(date, human_incidence, catt_incidence, goat_incidence) |> 
  filter(!is.na(catt_incidence)) |> 
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .)))

# Cases
df_1_trend_cases <- df_tot_cases_trend |>
  merge(df_tot_pop_trend, by = "date") |>
  filter(!is.na(date)) |>
  mutate(
    human_incidence = round((hum_cases / hum_pop) * 1000, 4),
    catt_incidence = round((catt_cases / catt_pop) * 1000000, 4),
    cam_incidence = round((cam_cases / cam_pop) * 1000000, 4),
    goat_incidence = round((goat_cases / goat_pop) * 1000000, 4),
    shp_incidence = round((shp_cases / sheep_pop) * 1000000, 4),
    date = as.Date(date)
  ) |>
  dplyr::select(date, contains(c("incidence", "cases"))) |>
  as_tibble() |> 
  mutate(
    
  )
write.csv(df_1_trend_cases, "individual_incidence_cases.csv")


## The differenced one
date <- df_1_trend$date[-1]

df_1_trend_diff <- df_1_trend |> 
  reframe(across(contains("incidence"), ~ diff(., na.rm = T))) |> 
  mutate(date = as.Date(date))

df_cum_trend <- df_tot_cases_trend |>
  merge(df_tot_pop_trend, by = "date") |>
  filter(!is.na(date)) |>
  rowwise() |> 
  mutate(animal_cases = sum(catt_cases, goat_cases, shp_cases, cam_cases, na.rm = T),
         animal_pop = sum(catt_pop, goat_pop, sheep_pop, cam_pop, na.rm = T),
        # animal_cases = ifelse(animal_cases == 0, NA, animal_cases),
         human_incidence = round((hum_cases / hum_pop) * 1000, 4),
         animal_incidence = round((animal_cases / animal_pop) * 1000000, 4)
  ) |> 
  dplyr::select(date, contains("incidence"))
write.csv(df_cum_trend, "combined_incidence.csv")

# Cases
df_cum_trend_cases <- df_tot_cases_trend |>
  merge(df_tot_pop_trend, by = "date") |>
  filter(!is.na(date)) |>
  rowwise() |> 
  mutate(animal_cases = sum(catt_cases, goat_cases, shp_cases, cam_cases, na.rm = T),
         animal_pop = sum(catt_pop, goat_pop, sheep_pop, cam_pop, na.rm = T),
        # animal_cases = ifelse(animal_cases == 0, NA, animal_cases),
         human_incidence = round((hum_cases / hum_pop) * 1000, 4),
         animal_incidence = round((animal_cases / animal_pop) * 1000000, 4)
  ) |> 
  dplyr::select(date, contains(c("incidence", "cases")))
write.csv(df_cum_trend_cases, "combined_incidence_cases.csv")

# The differenced one
df_cum_trend_diff <- df_cum_trend |> 
  arrange(date) |>  
  as.data.frame() |>
  reframe(across(c(human_incidence, animal_incidence), ~ diff(., 1, na.rm = T))) |> 
  mutate(date = as.Date(date))

trend_data <- df_1_trend %>%
  pivot_longer(cols = -date) %>%
  mutate(
    name = factor(name, levels = unique(name)),
    name = factor(name, labels = c(
      "Human Incidence", "Cattle Incidence", "Goat Incidence",
      "Sheep Incidence", "Camel Incidence"
    ))
  )

all_plus_hum <- df_cum_trend |> 
  #filter(!is.na(animal_incidence)) |> 
  mutate(animal_incidence = ifelse(is.na(animal_incidence), 0, animal_incidence)) |> 
  ggplot(aes(date)) +
  geom_point(aes(y = animal_incidence), col = "black", size = 1) +
  geom_line(aes(y = animal_incidence), col = "black", linewidth = 1) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(color = "black", hjust = 0.5, size = 12),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "bottom",
    legend.text = element_text(color = "black")
  ) +
  ylab("Animal Incidence") +
  ggtitle('Animal Incidence')
all_plus_hum

# All except humans
species_plt <- trend_data %>%
  filter(name != "Human Incidence") |>
  mutate(value = ifelse(is.na(value), 0, value)) |> 
  ggplot(aes(x = date)) +
  geom_line(aes(y = value, col = name), linewidth = 1) +
  geom_point(aes(y = value, col = name), size = 2) +
  theme_light() +
  #facet_wrap(~name, scales = "free", ncol = 3) +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(color = "black", hjust = 0.5, size = 12),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "bottom",
    legend.text = element_text(color = "black")
  ) +
  ylab("Incidence/1,000,000 population") +
  xlab("Year") +
  labs(col = "Species", title = "Incidence rate for cattle, goats, sheep and camels")
species_plt

# Humans
humans_plt <- trend_data %>%
  filter(name == "Human Incidence") |>
  mutate(value = ifelse(is.na(value), 0, value)) |> 
  ggplot(aes(x = date)) +
  geom_line(aes(y = value), linewidth = 1) +
  geom_point(aes(y = value), size = 2) +
  theme_light() +
  #facet_wrap(~name, scales = "free", ncol = 3) +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(color = "black", hjust = 0.5, size = 12),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "bottom",
    legend.text = element_text(color = "black")
  ) +
  ylab("Incidence/1,000 population") +
  xlab("Year") +
  labs(col = "Species", title = "Incidence rate for humans")
humans_plt

```

```{r testing for stationarity}

# At county Level
adf1 <- adf.test(na.omit(df_1$human_incidence))
adf2 <- adf.test(na.omit(df_1$catt_incidence))
adf3 <- adf.test(na.omit(df_1$cam_incidence))
adf4 <- adf.test(na.omit(df_1$goat_incidence))
adf5 <- adf.test(na.omit(df_1$shp_incidence))

adf_res <- data.frame(
  variable = c(
    "Human Incidence",
    "Cattle Incidence",
    "Camel Incidence",
    "Goat Incidence",
    "Sheep Incidence"
  ),
  statistic = c(
    adf1$statistic,
    adf2$statistic,
    adf3$statistic,
    adf4$statistic,
    adf5$statistic
  ),
  
  `P Value` = c(
    adf1$p.value,
    adf2$p.value,
    adf3$p.value,
    adf4$p.value,
    adf5$p.value
  )
  
  
) %>%
  mutate(across(where(is.numeric), ~round(., 3))) |> 
knitr::kable(
    align = "l",
    caption = "Results of Augmented Dickey-Fuller Test for Stationarity at County Level",
    format = "pipe",
    latex_options = "hold_position"
  )

# At National Level
adf_trend1 <- adf.test(na.omit(df_1_trend$human_incidence))
adf_trend2 <- adf.test(na.omit(df_1_trend$catt_incidence))
adf_trend3 <- adf.test(na.omit(df_1_trend$cam_incidence))
adf_trend4 <- adf.test(na.omit(df_1_trend$goat_incidence))
adf_trend5 <- adf.test(na.omit(df_1_trend$shp_incidence))

adf_trend_res <- data.frame(
  Variable = c(
    "Human Incidence",
    "Cattle Incidence",
    "Camel Incidence",
    "Goat Incidence",
    "Sheep Incidence"
  ),
  Statistic = c(
    adf_trend1$statistic,
    adf_trend2$statistic,
    adf_trend3$statistic,
    adf_trend4$statistic,
    adf_trend5$statistic
  ),
  
  `P Value` = c(
    adf_trend1$p.value,
    adf_trend2$p.value,
    adf_trend3$p.value,
    adf_trend4$p.value,
    adf_trend5$p.value
  )
  
  
) |>
  mutate(across(where(is.numeric), ~round(., 3))) |> 
  knitr::kable(
    align = "l",
    caption = "Results of Augmented Dickey-Fuller Test for Stationarity at National Level",
    format = "pipe",
    latex_options = "hold_position"
  )

## All incidences combined

adf_combined <- adf.test(na.omit(df_cum$animal_incidence))
adf_combined_trend <- adf.test(df_cum_trend$animal_incidence)


adf_combined_results <- data.frame(
  "Level" = c("County", "National"),
  "Statistic" = c(adf_combined$statistic, adf_combined_trend$statistic),
  "P Value" = c(adf_combined$p.value, adf_combined_trend$p.value)
)
```

```{r moving average smoothing}

# dplyr::select columns containing "incidence"
incidence_cols <- grep("incidence", names(df_1_trend), value = TRUE)

# Apply moving average smoothing to dplyr::selected columns
smoothed_df <- df_1_trend %>%
  mutate(across(incidence_cols, ~ifelse(is.na(.), 0, .))) |> 
  mutate(across(all_of(incidence_cols), ~zoo::rollmean(., k = 4, fill = NA), .names = "smoothed_{.col}")) |> 
  na.omit()

# Print the first few rows of the smoothed data
head(smoothed_df)

order <- c(
  "smoothed_human_incidence",
  "smoothed_catt_incidence",
  "smoothed_goat_incidence",
  "smoothed_shp_incidence",
  "smoothed_cam_incidence"
)


order2 <- c("Human Incidence", "Cattle Incidence", "Camel Incidence", "Goat Incidence", 
"Sheep Incidence")

trend_data_smoothed <- smoothed_df %>%
  dplyr::select(date, contains("smoothed")) %>%
  pivot_longer(cols = -date) %>%
  mutate(
    name = factor(name, levels = order),
    name = factor(name, labels = c(
      "Human Incidence", "Cattle Incidence", "Goat Incidence",
      "Sheep Incidence", "Camel Incidence"
    ))
  ) 

df_long <- df_1_trend %>%
  mutate(across(incidence_cols, ~ifelse(is.na(.), 0, .))) |> 
  dplyr::select(date, contains("incidence")) %>%
  pivot_longer(cols = -date) %>%
  mutate(
    name = case_when(
    name == "human_incidence" ~ "Human Incidence",
    name == "catt_incidence" ~ "Cattle Incidence",
    name == "goat_incidence" ~ "Goat Incidence",
    name == "shp_incidence" ~ "Sheep Incidence",
    TRUE ~ "Camel Incidence"
  )) |> 
  mutate(
    name = factor(name, levels = order2),
    name = factor(name, labels = c(
      "Human Incidence", "Cattle Incidence", "Goat Incidence",
      "Sheep Incidence", "Camel Incidence"
    ))
  )

species_sm_plt <- trend_data_smoothed |> 
  filter(name != "Human Incidence") |> 
  ggplot(aes(x = date)) +
  geom_line(aes(y = value, col = name), linewidth = 1) +
  #geom_point(aes(y = value, col = name)) +
  # facet_wrap(~name, scales = "free", ncol = 3) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(color = "black", hjust = 0.5, size = 12),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "bottom",
    legend.text = element_text(color = "black")
  ) +
  ylab("Incidence/1,000,000 population") +
  xlab("Year") +
  labs(col = "Species", title = "Smoothed Incidence rate for cattle, goats, sheep and camels")

species_sm_plt

# With points
species_sm_plt_points <- trend_data_smoothed |> 
  filter(name != "Human Incidence") |> 
  ggplot(aes(x = date)) +
  geom_line(aes(y = value, col = name), linewidth = 1) +
  geom_point(data = trend_data |> 
               mutate(value = ifelse(is.na(value), 0, value)) |> 
               filter(name != "Human Incidence"),
             
             aes(y = value, col = name), size = 1) +
  # facet_wrap(~name, scales = "free", ncol = 3) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(color = "black", hjust = 0.5, size = 12),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "bottom",
    legend.text = element_text(color = "black")
  ) +
  ylab("Incidence/1,000,000 population") +
  xlab("Year") +
  labs(col = "Species", title = "Smoothed Incidence rate for cattle, goats, sheep and camels")

species_sm_plt_points

humans_sm_plt <- trend_data_smoothed |> 
  filter(name == "Human Incidence") |> 
  ggplot(aes(x = date)) +
  geom_line(aes(y = value), linewidth = 1) +
  geom_point(data = df_1_trend, aes(y = human_incidence), col = "red") +
  # facet_wrap(~name, scales = "free", ncol = 3) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(color = "black", hjust = 0.5, size = 12),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "bottom",
    legend.text = element_text(color = "black")
  ) +
  ylab("Incidence/1,000 population") +
  xlab("Year") +
  labs(col = "Species", title = "Smoothed incidence rate for humans")
humans_sm_plt

humans_sm_plt_nopoints <- trend_data_smoothed |> 
  filter(name == "Human Incidence") |> 
  ggplot(aes(x = date)) +
  geom_line(aes(y = value), linewidth = 1) +
  #geom_point(data = df_1_trend, aes(y = human_incidence), col = "red") +
  # facet_wrap(~name, scales = "free", ncol = 3) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(color = "black", hjust = 0.5, size = 12),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "bottom",
    legend.text = element_text(color = "black")
  ) +
  ylab("Incidence/1,000 population") +
  xlab("Year") +
  labs(col = "Species", title = "Smoothed incidence rate for humans")
humans_sm_plt_nopoints

## For all animal incidence combined
df_cum_trend <- df_cum_trend |> 
  mutate(date = as.Date(date))

incidence_cols_cum <- grep("incidence", names(df_cum_trend), value = TRUE)

# Apply moving average smoothing to dplyr::selected columns
# dplyr::select columns containing "incidence"
incidence_cols <- grep("incidence", names(df_cum_trend), value = TRUE)

# Apply moving average smoothing to dplyr::selected columns
smoothed_df_combined <- df_cum_trend %>%
  mutate(animal_incidence = ifelse(is.na(animal_incidence), 0, animal_incidence)) |> 
  as_tibble() |>
  mutate(across(
    all_of(incidence_cols),
    ~ zoo::rollmean(., k = 2, fill = NA),
    .names = "smoothed_{.col}"
  )) |>
  na.omit()

animal_sm_plt <- smoothed_df_combined |> 
  ggplot(aes(x = date)) +
  geom_line(aes(y = smoothed_animal_incidence), linewidth = 1) +
  geom_point(data = df_cum_trend |> 
               mutate(animal_incidence = ifelse(is.na(animal_incidence), 0, animal_incidence)) 
               , aes(y = animal_incidence), col = "red") +
  # facet_wrap(~name, scales = "free", ncol = 3) +
  theme_light() +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(color = "black", hjust = 0.5, size = 12),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "bottom",
    legend.text = element_text(color = "black")
  ) +
  ylab("Incidence/1,000,000 population") +
  xlab("Year") +
  labs(col = "Species", title = "Smoothed incidence rate for animal incidence combined")
animal_sm_plt
```

```{r Descriptive statistics}

table1.1 <- df_incidence2.1 |>
  dplyr::select(county, diagnosis, contains("cases")) |>
  mutate(
    catt_cases = ifelse(catt_cases >= 69, round(mean(catt_cases, na.rm = T)), catt_cases),
    goat_cases = ifelse(goat_cases > 28, round(mean(goat_cases, na.rm = T)), goat_cases)
  ) |>
  pivot_longer(cols = -c(county, diagnosis)) %>%
  group_by(name) |>
  group_by(name, Diagnosis = diagnosis) |>
  summarise(Cases = sum(value, na.rm = T)) |> 
  mutate(
    Species = recode(
      name,
      "cam_cases" = 'Camels',
      "hum_cases" = 'Humans',
      "goat_cases" = 'Goats',
      "shp_cases" = "Sheep",
      "catt_cases" = "Cattle"
    ) 
  ) |> 
  ungroup() |> 
  dplyr::select(-name) |> 
  group_by(Species, Diagnosis) |> 
  group_by(Species) |> 
  mutate(`Percent(%)` = round((Cases/sum(Cases, na.rm = T)) * 100, 2)) |> 
  mutate(Cases = format(round(Cases), big.mark = ',') |> str_squish()) |> 
  dplyr::select(3, 1,2,4) |> 
  knitr::kable(align = "l", 
               caption = "Number of cases according to the type of Diagnosis", 
               format = "pipe",
               latex_options = "hold_position")
table1.1


write.csv( df_incidence |> 
  dplyr::select(county, diagnosis, contains("cases")) |> 
  pivot_longer(cols = -c(county, diagnosis)) %>%
  group_by(name) |> 
  group_by(name, Diagnosis = diagnosis) |> 
  summarise(Cases = sum(value, na.rm = T)) |> 
  mutate(
    Species = recode(
      name,
      "cam_cases" = 'Camels',
      "hum_cases" = 'Humans',
      "goat_cases" = 'Goats',
      "shp_cases" = "Sheep",
      "catt_cases" = "Cattle"
    ) 
  ) |> 
  ungroup() |> 
  dplyr::select(-name) |> 
  group_by(Species, Diagnosis) |> 
  group_by(Species) |> 
  mutate(`Percent(%)` = round((Cases/sum(Cases)) * 100, 2)) |> 
  dplyr::select(3, 1,2,4),
  
  "cases_table.csv")

# The descriptive statistics are for the Incidence Rate National Wide
table2 <- df_1 %>%
  dplyr::select(county, contains("incidence")) |>
  pivot_longer(cols = -1) %>%
  group_by(name) %>%
  summarise(
    `Mean Incidence Rate` = mean(value, na.rm = TRUE),
    minimum = min(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    max = max(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE)
  ) %>%
  arrange(desc(`Mean Incidence Rate`)) %>%
  mutate(
    name = case_when(
     name ==  "human_incidence" ~   "Human",
      name =="catt_incidence"  ~  "Cattle", 
      name == "goat_incidence" ~  "Goat", 
      name == "cam_incidence"  ~ "Camel", 
      name == "shp_incidence" ~  "Sheep"
     ),
    Cases = comma(`Mean Incidence Rate`),
    Minimum = comma(minimum),
    Median = comma(median),
    Maximum = comma(max),
    `Standard Deviation` = comma(sd)
  ) |> 
  dplyr::select(Species = name,
         `Mean Incidence Rate`,
         Minimum,
         Median,
         Maximum,
         `Standard Deviation`) |>
  knitr::kable(
    align = "l",
    caption = "Descriptive Statistics for Incidence Rate",
    format = "pipe",
    latex_options = "hold_position"
  )

table2

write.csv(df_1 %>%
  dplyr::select(county, contains("incidence")) |>
  pivot_longer(cols = -1) %>%
  group_by(name) %>%
  summarise(
    `Mean Incidence Rate` = mean(value, na.rm = TRUE),
    minimum = min(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    max = max(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE)
  ) %>%
  arrange(desc(`Mean Incidence Rate`)) %>%
  mutate(
    name = case_when(
     name ==  "human_incidence" ~   "Human",
      name =="catt_incidence"  ~  "Cattle", 
      name == "goat_incidence" ~  "Goat", 
      name == "cam_incidence"  ~ "Camel", 
      name == "shp_incidence" ~  "Sheep"
     ),
    Cases = comma(`Mean Incidence Rate`),
    Minimum = comma(minimum),
    Median = comma(median),
    Maximum = comma(max),
    `Standard Deviation` = comma(sd)
  ) |> 
  dplyr::select(Species = name,
         `Mean Incidence Rate`,
         Minimum,
         Median,
         Maximum,
         `Standard Deviation`),
  "descriptive_table.csv")
```

```{r The spatial analysis/plotting}

# Cases per year per county
df_tot_cases_spatial <- df_incidence2.1 |>
  group_by(year = year(as.Date(date)), county) |>
  summarise(across(contains("cases"), ~ sum(., na.rm = T))) |>
  mutate(across(contains('cases'), ~ ifelse(. == 0, NA, .)))

# Population per year, per county
df_pop_spatial <- df_incidence2.1 |>
  dplyr::select(date, county, contains("pop")) %>%
  distinct(.) |>
  as_tibble() |>
  group_by(year = year(as.Date(date)), county) %>%
  summarise(across(where(is.numeric), ~ unique(.))) |>
  mutate(across(contains('cases'), ~ ifelse(. == 0, NA, .)))

# Individual Cases per year per county month
df_tot_cases_spatial_month <- df_incidence2.1 |>
  group_by(date, county) |>
  summarise(across(contains("cases"), ~ sum(., na.rm = T))) |>
  mutate(year = year(as.Date(date))) |>
  merge(df_pop_spatial, by = c("year", "county")) |>
  filter(!is.na(year)) |>
  mutate(
    human_incidence = round((hum_cases / pop) * 1000, 4),
    catt_incidence = round((catt_cases / catt_pop) * 1000000, 4),
    cam_incidence = round((cam_cases / cam_pop) * 1000000, 4),
    goat_incidence = round((goat_cases / goat_pop) * 1000000, 4),
    shp_incidence = round((shp_cases / sheep_pop) * 1000000, 4)
  ) |>
  dplyr::select(year, date, county, contains(c("incidence", "cases"))) |>
  as_tibble()

write.csv(df_tot_cases_spatial_month,
          "df_tot_cases_spatial_month.csv")

df_tot_cases_spatial_month_complete <- df_tot_cases_spatial_month |>
  dplyr::select(
    year,
    date,
    county,
    human_incidence,
    goat_incidence,
    catt_incidence,
    goat_cases,
    catt_cases,
    hum_cases
  )

write.csv(df_tot_cases_spatial_month_complete,
          "df_tot_cases_spatial_month_complete.csv")

# Combined cases
df_spatial_cum_month <- df_tot_cases_spatial_month |>
  merge(df_pop_spatial, by = c("year", "county")) |>
  rowwise() |>
  mutate(
    animal_cases = sum(catt_cases, goat_cases, shp_cases, cam_cases, na.rm = T),
    animal_pop = sum(catt_pop, goat_pop, sheep_pop, cam_pop, na.rm = T) %>%
      ifelse(. == 0, NA, .),
    animal_incidence = round((animal_cases / animal_pop) * 1000000, 4),
    human_incidence = round((hum_cases / pop) * 1000, 4)
  ) |>
  dplyr::select(date, year, county, contains(c("incidence", "cases"))) |>
  as_tibble()
write.csv(df_spatial_cum_month, "df_spatial_cum_month.csv")

# Population per year, per county
df_pop_spatial <- df_incidence2.1 |>
  dplyr::select(date, county, contains("pop")) %>%
  distinct(.) |>
  as_tibble() |>
  group_by(year = year(date), county) %>%
  summarise(across(where(is.numeric), ~ unique(.))) |>
  mutate(across(contains('cases'), ~ ifelse(. == 0, NA, .)))

df_spatial <- df_tot_cases_spatial |>
  merge(df_pop_spatial, by = c("year", "county")) |>
  filter(!is.na(year)) |>
  mutate(
    human_incidence = round((hum_cases / pop) * 1000, 4),
    catt_incidence = round((catt_cases / catt_pop) * 1000000, 4),
    cam_incidence = round((cam_cases / cam_pop) * 1000000, 4),
    goat_incidence = round((goat_cases / goat_pop) * 1000000, 4),
    shp_incidence = round((shp_cases / sheep_pop) * 1000000, 4)
  ) |>
  dplyr::select(year, county, contains("incidence")) |>
  #mutate(across(is.numeric, ~ifelse(is.na(.), 0, .))) |>
  as_tibble()

df_spatial_cum <- df_tot_cases_spatial |>
  merge(df_pop_spatial, by = c("year", "county")) |>
  rowwise() |>
  mutate(
    animal_cases = sum(catt_cases, goat_cases, shp_cases, cam_cases, na.rm = T),
    animal_pop = sum(catt_pop, goat_pop, sheep_pop, cam_pop, na.rm = T) %>%
      ifelse(. == 0, NA, .),
    animal_cases = ifelse(animal_cases == 0, NA, animal_cases),
    animal_incidence = round((animal_cases / animal_pop) * 1000000, 4),
    human_incidence = round((hum_cases / pop) * 1000, 4)
  ) |>
  dplyr::select(year, county, contains("incidence")) |>
  as_tibble()

# Checking for mismatch of county names in the shapefiles and in our data
setdiff(shp$Name, df_spatial$county)
setdiff(df_spatial$county, shp$Name)

# Replacing Muranga to Murang'a
shp <- shp |>
  mutate(Name = ifelse(Name == "Muranga", "Murang'a", Name))
setdiff(shp$Name, df_spatial$county)
setdiff(df_spatial$county, shp$Name)
length(unique(df_spatial$county))

# Merging
df_spatial_merged <- df_spatial |>
  merge(shp, by.x = "county", by.y = 'Name')
indi_incidence_columns <- grep("incidence", names(df_spatial_merged), value = TRUE)

df_spatial_merged <- df_spatial_merged |>
  mutate(across(
    all_of(indi_incidence_columns),
    ~ cut(
      .,
      breaks = quantile(., na.rm = TRUE),
      include.lowest = TRUE
    ),
    .names = "{col}_range"
  )) |>
  st_as_sf()

df_spatial_merged_cum <- df_spatial_cum |>
  merge(shp, by.x = "county", by.y = 'Name')

all_incidence_columns <- grep("incidence", names(df_spatial_merged_cum), value = TRUE)
df_spatial_merged_cum <- df_spatial_merged_cum |>
  mutate(across(
    all_of(all_incidence_columns),
    ~ cut(
      .,
      breaks = quantile(., na.rm = TRUE),
      include.lowest = TRUE
    ),
    .names = "{col}_range"
  )) |>
  st_as_sf()

# Plotting
# Convert year to factor for better plotting
df_spatial_merged$year <- as.factor(df_spatial_merged$year)
df_spatial_merged_cum$year <- as.factor(df_spatial_merged_cum$year)

# All animals incidence
cate_animal <- length(levels(df_spatial_merged_cum$animal_incidence_range))
animals <- df_spatial_merged_cum |>
  mutate(animal_incidence_range = ifelse(
    is.na(animal_incidence_range),
    "0",
    as.character(animal_incidence_range)
  ) %>%
    factor(
      .,
      levels = c("0",  # These are the quantiles
                 "[0.0074,0.278]",
                 "(0.278,1.37]" ,
                 "(1.37,4.26]" ,
                 "(4.26,336]")
    )) |>
  ggplot() +
  geom_sf(aes(fill = animal_incidence_range)) +
  scale_fill_manual(
    values = c("white", brewer.pal(cate_animal, "YlOrRd")),
    labels = function(breaks) {
      str_replace_all(breaks, "\\[|\\)|\\]|\\(", "") %>% str_replace_all(., ",", " - ")
    },
    
    na.value = "white"
  ) +
  theme_void() +
  facet_wrap( ~ year, nrow = 1) +
  theme(
    plot.title = element_text(
      color = "black",
      hjust = .5,
      size = 16
    ),
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, colour = "black"),
    legend.key.size = unit(0.7, "cm"),
    strip.text = element_text(colour = "black", size = 16)
  ) +
  ggtitle("Animals") +
  labs(fill = "Animals")
animals

# Humans
cate_human <-
  length(levels(df_spatial_merged_cum$human_incidence_range))

humans <-  
  # mutate(human_incidence_range = ifelse(is.na(human_incidence_range),
  #                                        "0",
  #                                        as.character(human_incidence_range)) %>%
  #          factor(., levels = c(
  #            "0",
  #            "[0.113,0.865]",
  #            "(0.865,1.85]",
  #            "(1.85,5.62]",
  #            "(5.62,67.4]"
  #          ))) |>
  ggplot() +
  #geom_sf(data = shp, fill = 'white') +
  geom_sf(data = df_spatial_merged_cum, aes(fill = human_incidence_range)) +  
  scale_fill_manual(
    values = c(brewer.pal(cate_human, "YlOrRd")),
    labels =
      function(breaks) {
        str_replace_all(breaks, "\\[|\\)|\\]|\\(", "") %>% str_replace_all(., ",", " - ")
      },
    na.value = "white"
  ) +
  theme_void() +
  facet_wrap( ~ year, nrow = 1) +
  theme(
    plot.title = element_text(
      color = "black",
      hjust = .5,
      size = 16
    ),
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, colour = "black"),
    legend.key.size = unit(0.7, "cm"),
    strip.text = element_text(colour = "black", size = 16)
  ) +
  ggtitle("Humans") +
  labs(fill = "Humans")
humans

animals_humans <-
  wrap_plots(humans,
             animals,
             ncol = 1,
             guides = "keep") +
  plot_annotation(caption = "For humans, the incidence rate is per
                  1,000 population while for animal,
                  the incidence rate is per 1,000,000 population
                  ") &
  theme(plot.caption = element_text(size = 16, colour = "black"))

animals_humans

cate_catt <- length(levels(df_spatial_merged$catt_incidence_range))
cattle <- df_spatial_merged |>
  mutate(catt_incidence_range = ifelse(is.na(catt_incidence_range), 
                                         "0", 
                                         as.character(catt_incidence_range)) %>%
           factor(., levels = c(
             "0",
             "[0.0317,1.26]",
             "(1.26,2.94]",
             "(2.94,8.74]",
             "(8.74,235]" 
           ))) |> 
  ggplot() +
  geom_sf(aes(fill = catt_incidence_range)) +
  scale_fill_manual(values = c("white", brewer.pal(cate_catt, "YlOrRd")),
                        labels =
      function(breaks) {
        str_replace_all(breaks, "\\[|\\)|\\]|\\(", "") %>% str_replace_all(., ",", " - ")
      },
    
                    na.value = "white") +
  theme_void() +
  facet_wrap( ~ year, nrow = 1) +
  theme(
    plot.title = element_text(
      color = "black",
      hjust = .5,
      size = 16
    ),
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, colour = "black"),
    legend.key.size = unit(0.7, "cm"),
    strip.text = element_text(colour = "black", size = 16)
  ) +
  ggtitle("Cattle") +
  labs(fill = "Cattle") 
cattle

# Goats
cate_goat <- length(levels(df_spatial_merged$goat_incidence_range))
goat <- df_spatial_merged |>
  mutate(goat_incidence_range = ifelse(is.na(goat_incidence_range), 
                                         "0", 
                                         as.character(goat_incidence_range)) %>%
           factor(., levels = c(
             "0",
"[0.0178,0.259]",
"(0.259,1.09]",
"(1.09,6.98]" ,
"(6.98,690]" 
           ))) |> 
  ggplot() +
  geom_sf(aes(fill = goat_incidence_range)) +
  scale_fill_manual(values = c("white", brewer.pal(cate_goat, "YlOrRd")),
                        labels =
      function(breaks) {
        str_replace_all(breaks, "\\[|\\)|\\]|\\(", "") %>% str_replace_all(., ",", " - ")
      },
    
                    na.value = "white") +
  theme_void() +
  facet_wrap( ~ year, nrow = 1) +
  theme(
    plot.title = element_text(
      color = "black",
      hjust = .5,
      size = 16
    ),
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, colour = "black"),
    legend.key.size = unit(0.7, "cm"),
    strip.text = element_text(colour = "black", size = 16)
  ) +
  ggtitle("Goats") +
  labs(fill = "Goat") 
goat

# sheep
cate_shp <- length(levels(df_spatial_merged$shp_incidence_range))
sheep <- df_spatial_merged |>
  mutate(shp_incidence_range = ifelse(is.na(shp_incidence_range), 
                                         "0", 
                                         as.character(shp_incidence_range)) %>%
           factor(., levels = c(
             "0",
"[0.0259,0.13]",
"(0.13,1.46]",
"(1.46,3.85]",
"(3.85,45.1]" 
           ))) |> 
  ggplot() +
  geom_sf(aes(fill = shp_incidence_range)) +
  scale_fill_manual(values = c("white", brewer.pal(cate_shp, "YlOrRd")),
                        labels =
      function(breaks) {
        str_replace_all(breaks, "\\[|\\)|\\]|\\(", "") %>% str_replace_all(., ",", " - ")
      },
    
                    na.value = "white") +
  theme_void() +
  facet_wrap( ~ year, nrow = 1) +
  theme(
    plot.title = element_text(
      color = "black",
      hjust = .5,
      size = 16
    ),
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, colour = "black"),
    legend.key.size = unit(0.7, "cm"),
    strip.text = element_text(colour = "black", size = 16)
  ) +
  ggtitle("Sheep") +
  labs(fill = "Sheep") 
sheep

# Camels
cate_cam <- length(levels(df_spatial_merged$cam_incidence_range))
camels <- df_spatial_merged |>
  mutate(cam_incidence_range = ifelse(is.na(cam_incidence_range), 
                                         "0", 
                                         as.character(cam_incidence_range)) %>%
           factor(., levels = c(
             "0",
   "[0.613,0.97]",
   "(0.97,2.32]",
   "(2.32,3.36]",
   "(3.36,122]"  
           ))) |> 
  ggplot() +
  geom_sf(aes(fill = cam_incidence_range)) +
  scale_fill_manual(values = c("white", brewer.pal(cate_cam, "YlOrRd")),
                        labels =
      function(breaks) {
        str_replace_all(breaks, "\\[|\\)|\\]|\\(", "") %>% str_replace_all(., ",", " - ")
      },
    
                    na.value = "white") +
  theme_void() +
  facet_wrap( ~ year, nrow = 1) +
  theme(
    plot.title = element_text(
      color = "black",
      hjust = .5,
      size = 16
    ),
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, colour = "black"),
    legend.key.size = unit(0.7, "cm"),
    strip.text = element_text(colour = "black", size = 16)
  ) +
  ggtitle("Camels") +
  labs(fill = "Camels") 
camels

all_plots <-
  wrap_plots(humans,
             cattle,
             goat,
             sheep,
             camels,
             ncol = 1,
             guides = "keep") &
  theme(plot.caption = element_text(size = 16, colour = "black"))
dev.off()
ggsave("all_plots.png", dpi = 1e3, height = 18, width = 20)

```

```{r spatial plot, fig.height=18, fig.width=20}
all_plots

```

```{r Correlation}

# Correlation Plot
df_1_trend <- df_1_trend |> 
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .)))

cor_lag <- df_1_trend %>%
  as_tibble() %>%
  dplyr::select(-date) %>%
  setNames(c("Human",
             "Cattle",
             "Camel",
             "Goat",
             "Sheep")) |>
  cor() %>%
  ggcorrplot::ggcorrplot(type = "upper",
                         lab = TRUE,
                         lab_size = 6) +
  theme_light() +
  labs(subtitle = "Correlation between human incidence \nand animal incidences",
       x = NULL,
       y = NULL) +
  guides(fill = "none") +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(
      color = "black",
      hjust = 0.5,
      size = 35
    ),
    plot.subtitle = element_text(
      color = "black",
      hjust = 0.5,
      size = 14
    ),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "right",
    legend.text = element_text(color = "black")
  )
cor_lag

# Correlation Plot at lag 1
cor_lag1 <- df_1_trend %>%
  as_tibble() %>%
  mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
            list( ~ lag(., n = 1))) |>
  na.omit() |>
  dplyr::select(-date) %>%
  setNames(c("Human",
             "Cattle",
             "Camel",
             "Goat",
             "Sheep")) |>
  cor() %>%
  ggcorrplot::ggcorrplot(type = "upper",
                         lab = TRUE,
                         lab_size = 6) +
  theme_light() +
  labs(subtitle = "Correlation between\nhuman incidence \nand animal incidences at lag 1",
       x = NULL,
       y = NULL) +
  guides(fill = "none") +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(
      color = "black",
      hjust = 0.5,
      size = 20
    ),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "right",
    legend.text = element_text(color = "black")
  )
cor_lag1

# Correlation Plot at lag 2
cor_lag2 <- df_1_trend %>%
  as_tibble() %>%
  mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
            list( ~ lag(., n = 2))) |>
  na.omit() |>
  dplyr::select(-date) %>%
  setNames(c("Human",
             "Cattle",
             "Camel",
             "Goat",
             "Sheep")) |>
  cor() %>%
  ggcorrplot::ggcorrplot(type = "upper",
                         lab = TRUE,
                         lab_size = 6) +
  theme_light() +
  labs(subtitle = "Correlation between\nhuman incidence \nand animal incidences at lag 2",
       x = NULL,
       y = NULL) +
  guides(fill = "none") +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(
      color = "black",
      hjust = 0.5,
      size = 20
    ),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "right",
    legend.text = element_text(color = "black")
  )
cor_lag2

# Correlation Plot at lag 3
cor_lag3 <- df_1_trend %>%
  as_tibble() %>%
  mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
            list( ~ lag(., n = 3))) |>
  na.omit() |>
  dplyr::select(-date) %>%
  setNames(c("Human",
             "Cattle",
             "Camel",
             "Goat",
             "Sheep")) |>
  cor() %>%
  ggcorrplot::ggcorrplot(type = "upper",
                         lab = TRUE,
                         lab_size = 6) +
  theme_light() +
  labs(subtitle = "Correlation between\nhuman incidence \nand animal incidences at lag 3",
       x = NULL,
       y = NULL) +
  guides(fill = "none") +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(
      color = "black",
      hjust = 0.5,
      size = 20
    ),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "right",
    legend.text = element_text(color = "black")
  )
cor_lag3

# Correlation Plot at lag 4
cor_lag4 <- df_1_trend %>%
  as_tibble() %>%
  mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
            list( ~ lag(., n = 4))) |>
  na.omit() |>
  dplyr::select(-date) %>%
  setNames(c("Human",
             "Cattle",
             "Camel",
             "Goat",
             "Sheep")) |>
  cor() %>%
  ggcorrplot::ggcorrplot(type = "upper",
                         lab = TRUE,
                         lab_size = 6) +
  theme_light() +
  labs(subtitle = "Correlation between\nhuman incidence \nand animal incidences at lag 4",
       x = NULL,
       y = NULL) +
  guides(fill = "none") +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(
      color = "black",
      hjust = 0.5,
      size = 20
    ),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "right",
    legend.text = element_text(color = "black")
  )
cor_lag4

# Correlation Plot at lag 5
cor_lag5 <- df_1_trend %>%
  as_tibble() %>%
  mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
            list( ~ lag(., n = 5))) |>
  na.omit() |>
  dplyr::select(-date) %>%
  setNames(c("Human",
             "Cattle",
             "Camel",
             "Goat",
             "Sheep")) |>
  cor() %>%
  ggcorrplot::ggcorrplot(type = "upper",
                         lab = TRUE,
                         lab_size = 6) +
  theme_light() +
  labs(subtitle = "Correlation between\nhuman incidence \nand animal incidences at lag 5",
       x = NULL,
       y = NULL) +
  guides(fill = "none") +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(
      color = "black",
      hjust = 0.5,
      size = 20
    ),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "right",
    legend.text = element_text(color = "black")
  )
cor_lag5

# Correlation Plot at lag 6
cor_lag6 <- df_1_trend %>%
  as_tibble() %>%
  mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
            list( ~ lag(., n = 6))) |>
  na.omit() |>
  dplyr::select(-date) %>%
  setNames(c("Human",
             "Cattle",
             "Camel",
             "Goat",
             "Sheep")) |>
  cor() %>%
  ggcorrplot::ggcorrplot(type = "upper",
                         lab = TRUE,
                         lab_size = 6) +
  theme_light() +
  labs(subtitle = "Correlation between\nhuman incidence \nand animal incidences at lag 6",
       x = NULL,
       y = NULL) +
  theme(
    strip.background = element_rect(fill = "white", colour = "grey"),
    strip.text = element_text(color = "black", size = 12),
    axis.title = element_text(colour = "black"),
    axis.text = element_text(color = "black"),
    axis.ticks = element_line(color = "black", linewidth = 1),
    plot.title = element_text(
      color = "black",
      hjust = 0.5,
      size = 20
    ),
    axis.title.y = element_text(color = "black", size = 10),
    legend.position = "right",
    legend.text = element_text(color = "black")
  )
cor_lag6

all_cols <- wrap_plots(
  cor_lag,
  cor_lag1,
  cor_lag2,
  cor_lag3,
  cor_lag4,
  cor_lag5,
  cor_lag6,
  ncol = 3,
  guides = "collect"
) |> 
  plot_grid(
    rel_widths = c(7, 7,7)
  ) 
all_cols <- all_cols + theme(plot.title = element_text(size = 16),
                             axis.text.y = element_text(color = 'black', size = 13))

```

```{r lag with the highest average correlation}
lag_values <- 0:6  # Assuming you want lag values from 0 to 6
cor_dats <- list(

    cor_lag$data,
    cor_lag1$data,
    cor_lag2$data,
    cor_lag3$data,
    cor_lag4$data,
    cor_lag5$data,
    cor_lag6$data
  )
result_table <- tibble(
  Lag = lag_values,
  `Average Correlation` = cor_dats %>%
    map(~ filter(.x, Var1 == "Human")) %>%
    map_dbl(~mean(.$value))
) |>
  arrange(desc(`Average Correlation`)) |> 
  knitr::kable(align = "l", 
               caption = "Average correlation between human incidence and other species incidence, ordered from the largest to the smallest", 
               format = "pipe",
               latex_options = "hold_position")

print(result_table)
 # This helps us to choose the lag with the highest average correlation

```

```{r Models}

## This model fits the data without differencing, at difference lags, (0-3) and for individual 
## animal incidences.

df_1_trend <- df_1_trend |> 
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .)))

# Models with NA
run_lag_models <- function(df, max_lag = 3, ...) {
  suppressMessages({
  result_df <- tibble()
  
  for (lag_value in 0:max_lag) {
    df_lagged <- df |>
      as_tibble() %>%
      mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
                list( ~ lag(., n = lag_value))) |>
      na.omit() |>
      mutate(date = as.Date(date))
    
    mod <- df_lagged |>
      as_tsibble() |>
      model(
        TSLM(
          (human_incidence) ~ cam_incidence + shp_incidence + catt_incidence + goat_incidence
        )
      ) |>
      report()
    
    mod_results <- tidy(mod) %>%
      dplyr::select(-.model) %>%
      as_tibble() %>%
      mutate(term = case_when(
        term == "goat_incidence" ~ "Goat Incidence",
        term == "catt_incidence" ~ "Cattle incidence",
        term == "shp_incidence" ~ "Sheep incidence",
        term == "cam_incidence" ~ "Camel incidence",
        TRUE ~ as.character(term) 
      ),
      variable = term
      ) |>  dplyr::select(6, 2:5) |> 
      group_by(variable) %>%
      mutate(
        conf_low = min(estimate - std.error * 1.645),
        conf_high = max(estimate + std.error * 1.645)
      ) %>%
      mutate(lag = lag_value)
    
    adj_r_squared <- glance(mod) %>%
      dplyr::select(r_squared, AIC, adj_r_squared)
    
    mod_results <- bind_cols(mod_results, adj_r_squared) |>
      mutate(across(c(estimate, std.error, statistic, p.value, conf_low, conf_high, adj_r_squared), ~round(., 3))) |>
      mutate(significance = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) |> 
      dplyr::select(c(1:8, 12, 9:11))
      
    result_df <- bind_rows(result_df, mod_results)
  }
  })
  return(result_df)
}

# Models with complete cases
complete_run_lag_models <- function(df, max_lag = 3, ...) {
  suppressMessages({
  result_df <- tibble()
  
  for (lag_value in 0:max_lag) {
    df_lagged <- df |>
      as_tibble() %>%
      mutate_at(vars(catt_incidence, goat_incidence),
                list( ~ lag(., n = lag_value))) |>
      na.omit() |>
      mutate(date = as.Date(date))
    
    mod <- df_lagged |>
      as_tsibble() |>
      model(
        TSLM(
          human_incidence ~ catt_incidence + goat_incidence
        )
      ) |>
      report()
    
    mod_results <- tidy(mod) %>%
      dplyr::select(-.model) %>%
      as_tibble() %>%
      mutate(term = case_when(
        term == "goat_incidence" ~ "Goat Incidence",
        term == "catt_incidence" ~ "Cattle incidence",
        TRUE ~ as.character(term) 
      ),
      variable = term
      ) |>  dplyr::select(6, 2:5) |> 
      group_by(variable) %>%
      mutate(
        conf_low = min(estimate - std.error * 1.645),
        conf_high = max(estimate + std.error * 1.645)
      ) %>%
      mutate(lag = lag_value)
    
    adj_r_squared <- glance(mod) %>%
      dplyr::select(r_squared, AIC, adj_r_squared)
    
    mod_results <- bind_cols(mod_results, adj_r_squared) |>
      mutate(across(c(estimate, std.error, statistic, p.value, conf_low, conf_high, adj_r_squared), ~round(., 3))) |>
      mutate(significance = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) |> 
      dplyr::select(c(1:8, 12, 9:11))
      
    result_df <- bind_rows(result_df, mod_results)
  }
  })
  return(result_df)
}

non_diff_indivi_with_NA <- run_lag_models(df_1_trend)
  
non_diff_indivi_without_NA <-
  complete_run_lag_models(df_1_trend_complete)

diff_indivi <- run_lag_models(df_1_trend_diff |> 
                                mutate(across(contains('incidence'), ~ifelse(is.na(.), 0, .))))

## individual animal spp national level ------------------
write_csv(non_diff_indivi_with_NA |>
  dplyr::select(-r_squared, -AIC, -adj_r_squared), "non_diff_individual_with_NA.csv") 


write_csv(non_diff_indivi_without_NA |>
  dplyr::select(-r_squared, -AIC, -adj_r_squared), "non_diff_individual_without_NA.csv")

write_csv(diff_indivi, "diff_individual.csv")


## The following function fits the model, without differencing, at difference lags, (0-3) and for
## animal incidences combined

lag_models_full <- function(df, max_lag = 3, ...) {
  result_df <- tibble()
  
  for (lag_value in 0:max_lag) {
    df_lagged <- df |>
      as_tibble() %>%
      mutate_at(vars(animal_incidence),
                list(~ lag(., n = lag_value))) |>
      na.omit() |>
      mutate(date = as.Date(date))
    
    mod <- df_lagged |>
      as_tsibble() |>
      model(TSLM((human_incidence) ~ animal_incidence)) |>
      report()
    
    mod_results <- tidy(mod) %>%
      dplyr::select(-.model) %>%
      as_tibble() %>%
      mutate(
        term = case_when(
          term == "animal_incidence" ~ "Animal Incidence",
          TRUE ~ as.character(term)
        ),
        variable = term
      ) |>  dplyr::select(6, 2:5) |>
      group_by(variable) %>%
      mutate(
        conf_low = min(estimate - std.error * 1.645),
        conf_high = max(estimate + std.error * 1.645)
      ) %>%
      mutate(lag = lag_value)
    
    adj_r_squared <- glance(mod) |>
      dplyr::select(r_squared, adj_r_squared, AIC) |>
      as.data.frame()
    
    
    mod_results <- bind_cols(mod_results, adj_r_squared) |>
      mutate(across(
        c(
          estimate,
          std.error,
          statistic,
          p.value,
          conf_low,
          conf_high,
          adj_r_squared
        ),
        ~ round(., 3)
      )) |>
      mutate(significance = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) 
    
    result_df <- bind_rows(result_df, mod_results)
    
  }
  
  return(result_df)
}

non_diff_full_with_NA <- lag_models_full(df_cum_trend |> 
                                 mutate(across(contains('incidence'), ~ifelse(is.na(.), 0, .))))

non_diff_full_without_NA <- lag_models_full(df = df_cum_trend)

diff_full <- lag_models_full(df_cum_trend_diff)


## individual animal spp national level ---------
write.csv(non_diff_full_with_NA |>
  dplyr::select(-r_squared, -AIC, -adj_r_squared), "non_diff_full_with_NA.csv")

write.csv(non_diff_full_without_NA |>
  dplyr::select(-r_squared, -AIC, -adj_r_squared), "non_diff_full_without_NA.csv")

# write_csv(non_diff_full, "non_diff_full.csv")
# write_csv(diff_full, "diff_full.csv")

## Getting the AIC, R-Squared and Adjusted R squared for each lag
Table_lag_indivi_withoutNA <- non_diff_indivi_without_NA |> 
  dplyr::select(Lag = lag, `R-Squared(%)` = r_squared, `Adjusted R-Squared(%)` = adj_r_squared, AIC) |> 
  mutate(across(c(`R-Squared(%)` , `Adjusted R-Squared(%)`, ), ~ (. * 100))) |> 
    mutate(across(where(is.numeric), ~round(., 2))) |> 
  unique() |> 
  arrange(desc(`Adjusted R-Squared(%)`)) |> 
  knitr::kable(
    align = "l",
    caption = "The AIC, R-Squared and Adjusted R-Squared for each lag for individual species. The data as been arranged in decreasing order of Adjusted R-Squared",
    format = "pipe",
    latex_options = "hold_position"
  )
Table_lag_indivi_withoutNA


Table_lag_full <- non_diff_full_without_NA |> 
  dplyr::select(Lag = lag, `R-Squared(%)` = r_squared, `Adjusted R-Squared(%)` = adj_r_squared, AIC) |> 
  mutate(across(c(`R-Squared(%)` , `Adjusted R-Squared(%)`, ), ~ (. * 100))) |> 
    mutate(across(where(is.numeric), ~round(., 2))) |> 
  unique() |> 
  arrange(desc(`Adjusted R-Squared(%)`)) |> 
    knitr::kable(
    align = "l",
    caption = "The AIC, R-Squared and Adjusted R-Squared for each lag for combined species. The data as been arranged in decreasing order of Adjusted R-Squared",
    format = "pipe",
    latex_options = "hold_position"
  )
Table_lag_full
```

```{r forest plots}


# At lag 0
lag0_forest_df_lag0 <- non_diff_indivi_with_NA |>
  dplyr::select(-r_squared, -AIC, -adj_r_squared) %>% 
  filter(lag == 0 & variable != '(Intercept)')

lag0_forest_plot_lag0 <- lag0_forest_df_lag0 %>% 
  ggplot(aes(x = variable, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2, color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  
  coord_flip() +
  labs(
    #title = "Incidence Rate Ratios (IRRs) and Confidence Intervals",
    x = "Variable at lag 0",
    y = "Estimate"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_text(size = 12, color = 'black'),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "none"
  )+
  ggtitle('Individual animal incidence at Lag 0')




# At lag 3
lag3_forest_df_lag3 <- non_diff_indivi_with_NA |>
  dplyr::select(-r_squared, -AIC, -adj_r_squared) %>% 
  filter(lag == 3 & variable != '(Intercept)')

lag3_forest_plot_lag3 <- lag3_forest_df_lag3 %>% 
  ggplot(aes(x = variable, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2, color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  
  coord_flip() +
  labs(
    #title = "Incidence Rate Ratios (IRRs) and Confidence Intervals",
    x = "Variable at lag 3",
    y = "Estimate"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_text(size = 12, color = 'black'),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "none"
  ) +
  ggtitle('Individual animal incidence at Lag 3')


all_forest_individual <- lag0_forest_plot_lag0 +
lag3_forest_plot_lag3

dev.off()
ggsave(
  "images/all_forest_individual.png",
  width = 13,
  height = 6,
  dpi = 1e3
)


```

```{r forest plot combined}
# At lag 0
lag0_forest_df1_lag0 <- non_diff_full_with_NA |>
  dplyr::select(-r_squared, -AIC, -adj_r_squared) %>% 
  filter(lag == 0 & variable != '(Intercept)')

lag0_forest_plot1_lag0 <- lag0_forest_df1_lag0 %>% 
  ggplot(aes(x = variable, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2, color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  
  coord_flip() +
  labs(
    #title = "Incidence Rate Ratios (IRRs) and Confidence Intervals",
    x = "Variable at lag 0",
    y = "Estimate"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_text(size = 12, color = 'black'),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "none"
  )+
  ggtitle('Combined animal incidence at Lag 0')




# At lag 3
lag3_forest_df1_lag3 <- non_diff_full_with_NA |>
  dplyr::select(-r_squared, -AIC, -adj_r_squared) %>% 
  filter(lag == 3 & variable != '(Intercept)')

lag3_forest_plot1_lag3 <- lag3_forest_df1_lag3 %>% 
  ggplot(aes(x = variable, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2, color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  
  coord_flip() +
  labs(
    #title = "Incidence Rate Ratios (IRRs) and Confidence Intervals",
    x = "Variable at lag 3",
    y = "Estimate"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_text(size = 12, color = 'black'),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "none"
  ) +
  ggtitle('Combined animal incidence at Lag 3')


all_forest_comb <- lag0_forest_plot1_lag0 +
lag3_forest_plot1_lag3

dev.off()
ggsave(
  "images/all_forest_full.png",
  width = 13,
  height = 6,
  dpi = 1e3
)


# One plot
all_forests <- all_forest_individual/all_forest_comb

dev.off()
ggsave(
  "images/all_forest_full.png",
  width = 18,
  height = 10,
  dpi = 1e3
)


```

```{r All counties model (individual with NA) at lag 2 and 3}

# Lag 0
df_2_0 <- df_1 |> 
  as_tibble() %>%
  mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
            list( ~ dplyr::lag(., n = 0))) |>
  na.omit() |> 
  mutate(date = as.Date(date))

# Lag 3
df_2_3 <- df_1 |> 
  as_tibble() %>%
  mutate_at(vars(catt_incidence, cam_incidence, goat_incidence, shp_incidence),
            list( ~ dplyr::lag(., n = 3))) |>
  na.omit() |> 
  mutate(date = as.Date(date))


fit_county_model <- function(county_name, data, type) {
  # Subset the data for the specific county
  county_data <- filter(data, county == county_name)
  
  if (type == "full")
  {
    # Check if all incidences are zero
    if (all(county_data$animal_incidence == 0)) {
      message(paste("Skipping model for", county_name, "as all incidences are zero."))
      return(NULL)
    }
    
    # Fit the model
    mod_county <- county_data |>
      as_tsibble() |>
      model(
        TSLM(
          human_incidence ~ animal_incidence
        )
      ) |>
      tidy() |>
      dplyr::select(-.model) |>
      as_tibble() |>
      mutate(term = case_when(
        term == "animal_incidence" ~ "Animal Incidence",
        TRUE ~ as.character(term) 
      ),
      variable = term
      ) |>  
      dplyr::select(6, 2:5) |> 
      group_by(variable) %>%
      mutate(
        conf_low = min(estimate - std.error * 1.645),
        conf_high = max(estimate + std.error * 1.645)
      )
  }
  else if(type == "individual") {
  # Check if all incidences are zero
  if (all(county_data$catt_incidence == 0 &
          county_data$cam_incidence == 0 &
          county_data$goat_incidence == 0 &
          county_data$shp_incidence == 0)) {
    message(paste("Skipping model for", county_name, "as all incidences are zero."))
    return(NULL)
  }
  
  # Fit the model
  mod_county <- county_data |>
    as_tsibble() |>
    model(
      TSLM(
        human_incidence ~ catt_incidence + goat_incidence + shp_incidence +cam_incidence
      )
    ) |>
    tidy() |>
    dplyr::select(-.model) |>
    as_tibble() |>
    mutate(term = case_when(
      term == "goat_incidence" ~ "Goat Incidence",
      term == "catt_incidence" ~ "Cattle incidence",
      term == "shp_incidence" ~ "Sheep incidence",
      term == "cam_incidence" ~ "Camel incidence",
      TRUE ~ as.character(term) 
    ),
    variable = term
    ) |>  
    dplyr::select(6, 2:5) |> 
    group_by(variable) %>%
    mutate(
      conf_low = min(estimate - std.error * 1.645),
      conf_high = max(estimate + std.error * 1.645)
    )
  }
  return(mod_county)
}

county_names <- unique(df_2_0$county)

# Create a list to store the models for each county
models_list <- list()

# Initialize the data frame for each county
coefficients_df_lag0 <- data.frame(county = character(), 
                              variable = character(),
                              estimate = numeric(),
                              stringsAsFactors = FALSE)

# At lag 0
for (county_name in county_names) {
  message(paste("Fitting model for", county_name))
  

  
  # Fit the model
  mod_county <- fit_county_model(county_name, df_2_0, type = "individual")
  
  # Check if model fitting was successful
  if (!is.null(mod_county)) {
    # Extract coefficients, round off, and add to the data frame
    coefficients_df_lag0 <- bind_rows(coefficients_df_lag0, 
                                 mod_county %>% 
                                   mutate(county = county_name,
                                          estimate = round(estimate, 3)))
  }
}

coefficients_df_lag0 <- coefficients_df_lag0 |>
  mutate(significant = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) |> 
  mutate(across(c(3:8), ~round(., 3)))


lago_county_significant_individual <- coefficients_df_lag0 %>% 
  filter(significant == 'Significant' & variable != '(Intercept)') 

write_csv(lago_county_significant_individual, "lago_county_significant_individual.csv")

write_csv(coefficients_df_lag0, "individual_animal_incidence_per_county_lag0.csv")

# At lag 3
coefficients_df_lag3 <- data.frame(county = character(), 
                                   variable = character(),
                                   estimate = numeric(),
                                   stringsAsFactors = FALSE)

for (county_name in county_names) {
  message(paste("Fitting model for", county_name))
  
  
  
  # Fit the model
  mod_county <- fit_county_model(county_name, df_2_3, type = "individual")
  
  # Check if model fitting was successful
  if (!is.null(mod_county)) {
    # Extract coefficients, round off, and add to the data frame
    coefficients_df_lag3 <- bind_rows(coefficients_df_lag3, 
                                 mod_county %>% 
                                   mutate(county = county_name,
                                          estimate = round(estimate, 3)))
  }
}

coefficients_df_lag3 <- coefficients_df_lag3 |>
  mutate(significant = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) |> 
  mutate(across(c(3:8), ~round(., 3)))

lag3_county_significant_individual <- coefficients_df_lag3 %>% 
  filter(significant == 'Significant' & variable != '(Intercept)') 

write_csv(lag3_county_significant_individual, "lag3_county_significant_individual.csv")

write_csv(coefficients_df_lag3, "individual_animal_incidence_per_county_lag3.csv")

```

```{r forest plots at individual plot}
# At lag 0
indivi_lag0_df <- coefficients_df_lag0 %>% 
  filter( variable != '(Intercept)' & !is.nan(conf_low) & !is.na(conf_low)) %>% 
  filter(conf_high < 10 & conf_low > - 10)


indivi_lag0_plot <- indivi_lag0_df %>% 
  #filter(!county %in% c('West Pokot', 'Nandi', 'Elgeyo Marakwet')) %>% 
  ggplot(aes(x = county, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  
  coord_flip() +
  facet_wrap(~variable, scales = 'free') +
  labs(
    #title = "Incidence Rate Ratios (IRRs) and Confidence Intervals",
    x = "Variable at lag 0",
    y = "Estimate"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_text(size = 12, color = 'black'),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "none",
    strip.text = element_text(color = 'black', hjust = .5, face = 'bold', size = 13)
  ) +
  ggtitle('Individual animal incidence at Lag 0')
indivi_lag0_plot

dev.off()
ggsave(
  "images/indivi_lag0_forest.png",
  width = 18,
  height = 10,
  dpi = 1e3
)



# At lag 3
indivi_lag3_df <- coefficients_df_lag3 %>% 
  filter( variable != '(Intercept)' & !is.nan(conf_low) & !is.na(conf_low)) %>% 
  filter(conf_high < 10 & conf_low > - 10)


indivi_lag3_plot <- indivi_lag3_df %>% 
  ggplot(aes(x = county, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  
  coord_flip() +
  facet_wrap(~variable, scales = 'free') +
  labs(
    #title = "Incidence Rate Ratios (IRRs) and Confidence Intervals",
    x = "Variable at lag 3",
    y = "Estimate"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_text(size = 12, color = 'black'),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "none",
    strip.text = element_text(color = 'black', hjust = .5, face = 'bold', size = 13)
  ) +
  ggtitle('Individual animal incidence at Lag 3')
indivi_lag3_plot

dev.off()
ggsave(
  "images/indivi_lag3_forest.png",
  width = 18,
  height = 10,
  dpi = 1e3
)
```


```{r All counties model (combined with NA) at 0 and 3}
# Lag 2
df_cum_2_0 <- df_cum |> 
  as_tibble() %>%
  mutate_at(vars(animal_incidence),
            list( ~ dplyr::lag(., n = 0))) |>
  na.omit() |>
  mutate(date = as.Date(date))

# Lag 3
df_cum_2_3 <- df_cum |> 
  as_tibble() %>%
  mutate_at(vars(animal_incidence),
            list( ~ dplyr::lag(., n = 3))) |>
  na.omit() |>
  mutate(date = as.Date(date))

# At lag 0
county_names2_0 <- unique(df_cum_2_0$county)

# Create a list to store the models for each county
models_list <- list()

# Initialize the data frame for each county
coefficients_df2_lag0 <- data.frame(county = character(), 
                              variable = character(),
                              estimate = numeric(),
                              stringsAsFactors = FALSE)

for (county_name in county_names2_0) {
  message(paste("Fitting model for", county_name))
  
  
  
  # Fit the model
  mod_county <- fit_county_model(county_name, df_cum_2_0, type = "full")
  
  # Check if model fitting was successful
  if (!is.null(mod_county)) {
    # Extract coefficients, round off, and add to the data frame
    coefficients_df2_lag0 <- bind_rows(coefficients_df2_lag0, 
                                 mod_county %>% 
                                   mutate(county = county_name,
                                          estimate = round(estimate, 3)))
  }
}

coefficients_df2_lag0 <- coefficients_df2_lag0 |>
  mutate(across(c(3:8), ~round(., 3))) |> 
  mutate(significant = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) |> 
  as_tibble()

lag0_county_significant_combined <- coefficients_df2_lag0 %>% 
  filter(significant == 'Significant' & variable != '(Intercept)') 

write_csv(lag0_county_significant_combined,
          "lag0_county_significant_combined.csv")


write_csv(coefficients_df2_lag0, "all_animal_incidence_per_county_lag0.csv")

# At lag 3
county_names2_3 <- unique(df_cum_2_3$county)

# Create a list to store the models for each county
models_list <- list()

# Initialize the data frame for each county
coefficients_df2_lag3 <- data.frame(county = character(), 
                                    variable = character(),
                                    estimate = numeric(),
                                    stringsAsFactors = FALSE)

for (county_name in county_names2_3) {
  message(paste("Fitting model for", county_name))
  
  
  
  # Fit the model
  mod_county <- fit_county_model(county_name, df_cum_2_3, type = "full")
  
  # Check if model fitting was successful
  if (!is.null(mod_county)) {
    # Extract coefficients, round off, and add to the data frame
    coefficients_df2_lag3 <- bind_rows(coefficients_df2_lag3, 
                                       mod_county %>% 
                                         mutate(county = county_name,
                                                estimate = round(estimate, 3)))
  }
}

coefficients_df2_lag3 <- coefficients_df2_lag3 |>
  mutate(across(c(3:8), ~round(., 3))) |> 
  mutate(significant = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) |> 
  as_tibble()

lag3_county_significant_combined <- coefficients_df2_lag3 %>% 
  filter(significant == 'Significant' & variable != '(Intercept)') 

write_csv(lag3_county_significant_combined,
          "lag3_county_significant_combined.csv")

write_csv(coefficients_df2_lag3, "all_animal_incidence_per_county_lag3.csv")
```

```{r forest plots for combined incidence}

# At lag 0
comb_lag0_df <- coefficients_df2_lag0 %>% 
  filter( variable != '(Intercept)' & !is.nan(conf_low) & !is.na(conf_low)) 


comb_lag0_plot <- comb_lag0_df %>% 
  filter(county != 'Isiolo') %>% 
  ggplot(aes(x = county, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  
  coord_flip() +
  labs(
    #title = "Incidence Rate Ratios (IRRs) and Confidence Intervals",
    x = "Variable at lag 0",
    y = "Estimate"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_text(size = 12, color = 'black'),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "none"
  ) +
  ggtitle('Combined animal incidence at Lag 0')

# At lag 3
comb_lag3_df <- coefficients_df2_lag3 %>% 
  filter( variable != '(Intercept)' & !is.nan(conf_low) & !is.na(conf_low)) 


comb_lag3_plot <- comb_lag3_df %>% 
  filter(county != 'Isiolo') %>% 
  ggplot(aes(x = county, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  
  coord_flip() +
  labs(
    #title = "Incidence Rate Ratios (IRRs) and Confidence Intervals",
    x = "Variable at lag 3",
    y = "Estimate"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10, color = 'black'),
    axis.title.x = element_text(size = 12, color = 'black'),
    axis.title.y = element_text(size = 12),
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    legend.position = "none"
  ) +
  ggtitle('Combined animal incidence at Lag 3')

combined_forest <- comb_lag0_plot + comb_lag3_plot


dev.off()
ggsave(
  "images/combined_forest.png",
  width = 18,
  height = 10,
  dpi = 1e3
)

```


```{r # All counties model (individual without NA-complete case) at lag 2 and 3}
# Lag 2
df_2_0_complete <- df_1_complete |>
  as_tibble() %>%
  mutate_at(
    vars(catt_incidence, goat_incidence),
    list(~ dplyr::lag(., n = 2))
  ) |>
  na.omit() |>
  mutate(date = as.Date(date))

# Lag 3
df_2_3_complete  <- df_1_complete  |>
  as_tibble() %>%
  mutate_at(
    vars(catt_incidence, goat_incidence),
    list(~ dplyr::lag(., n = 4))
  ) |>
  na.omit() |>
  mutate(date = as.Date(date))


fit_county_model_complete <- function(county_name, data, type) {
  # Subset the data for the specific county
  county_data <- filter(data, county == county_name)
  
  if (type == "full")
  {
    # Check if all incidences are zero
    if (all(county_data$animal_incidence == 0)) {
      message(paste("Skipping model for", county_name, "as all incidences are zero."))
      return(NULL)
    }
    
    # Fit the model
    mod_county <- county_data |>
      as_tsibble() |>
      model(
        TSLM(
          human_incidence ~ animal_incidence
        )
      ) |>
      tidy() |>
      dplyr::select(-.model) |>
      as_tibble() |>
      mutate(term = case_when(
        term == "animal_incidence" ~ "Animal Incidence",
        TRUE ~ as.character(term) 
      ),
      variable = term
      ) |>  
      dplyr::select(6, 2:5) |> 
      group_by(variable) %>%
      mutate(
        conf_low = min(estimate - std.error * 1.645),
        conf_high = max(estimate + std.error * 1.645)
      )
  }
  else if (type == "individual") {
  # Check if all incidences are zero
  if (all(county_data$catt_incidence == 0 &
          county_data$goat_incidence == 0 )) {
    message(paste("Skipping model for", county_name, "as all incidences are zero."))
    return(NULL)
  }
  
  # Fit the model
  mod_county <- county_data |>
    as_tsibble() |>
    model(
      TSLM(
        human_incidence ~ catt_incidence + goat_incidence
      )
    ) |>
    tidy() |>
    dplyr::select(-.model) |>
    as_tibble() |>
    mutate(term = case_when(
      term == "goat_incidence" ~ "Goat Incidence",
      term == "catt_incidence" ~ "Cattle incidence",
      TRUE ~ as.character(term) 
    ),
    variable = term
    ) |>  
    dplyr::select(6, 2:5) |> 
    group_by(variable) %>%
    mutate(
      conf_low = min(estimate - std.error * 1.645),
      conf_high = max(estimate + std.error * 1.645)
    )
  }
  return(mod_county)
}

county_names <- unique(df_2_0_complete$county)

# Create a list to store the models for each county
models_list_withouNA <- list()

# Initialize the data frame for each county
coefficients_df_lag2_withoutNA <- data.frame(county = character(), 
                              variable = character(),
                              estimate = numeric(),
                              stringsAsFactors = FALSE)

# At lag 2
for (county_name in county_names) {
  message(paste("Fitting model for", county_name))
  

  
  # Fit the model
  mod_county <- fit_county_model_complete(county_name, df_2_0_complete, type = "individual")
  
  # Check if model fitting was successful
  if (!is.null(mod_county)) {
    # Extract coefficients, round off, and add to the data frame
    coefficients_df_lag2_withoutNA <- bind_rows(coefficients_df_lag2_withoutNA, 
                                 mod_county %>% 
                                   mutate(county = county_name,
                                          estimate = round(estimate, 3)))
  }
}

coefficients_df_lag2_withoutNA <- coefficients_df_lag2_withoutNA |>
  mutate(significant = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) |> 
  mutate(across(c(3:8), ~round(., 3)))
write_csv(coefficients_df_lag2_withoutNA, "individual_animal_incidence_per_county_lag2_withoutNA.csv")

# At lag 3
coefficients_df_lag3_withoutNA <- data.frame(county = character(), 
                                   variable = character(),
                                   estimate = numeric(),
                                   stringsAsFactors = FALSE)

for (county_name in county_names) {
  message(paste("Fitting model for", county_name))
  
  
  
  # Fit the model
  mod_county <- fit_county_model_complete(county_name, df_2_3_complete, type = "individual")
  
  # Check if model fitting was successful
  if (!is.null(mod_county)) {
    # Extract coefficients, round off, and add to the data frame
    coefficients_df_lag3_withoutNA <- bind_rows(coefficients_df_lag3_withoutNA, 
                                 mod_county %>% 
                                   mutate(county = county_name,
                                          estimate = round(estimate, 3)))
  }
}

coefficients_df_lag3_withoutNA <- coefficients_df_lag3_withoutNA |>
  mutate(significant = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) |> 
  mutate(across(c(3:8), ~round(., 3)))
write_csv(coefficients_df_lag3_withoutNA, "individual_animal_incidence_per_county_lag4_withoutNA.csv")

```

Note: All counties model (combined without NA-complete case) at lag 2 and 3 was omitted as it had no difference to combined with NA

# Mixed Effect Models

The section includes the mixed effect modelling. The mixed effect was done for both the incidence (as integers), cases and the proportion, and at lags 0 to 3.

```{r importing the data and packages for mixed effect modelling}

# Importing packages
if (require(pacman))
{
  p_load(
    tidyverse,
    tseries,
    data.table,
    scales,
    zoo,
    forecast,
    sf,
    patchwork,
    grid,
    fable,
    patchwork,
    xts,
    feasts,
    cowplot,
    broom,
    kableExtra,
    readxl,
    stringi,
    stringr,
    rKenyaCensus,
    knitr,
    purrr,
    RColorBrewer,
    tscount,
    lme4
  )
}

# Importing data
individual <- fread("individual_incidence_cases.csv")
combined <- fread("combined_incidence_cases.csv")
individual_county <- fread("df_tot_cases_spatial_month.csv")
combined_county <- fread("df_spatial_cum_month.csv")


# Complete
county_indivi_complete <- fread("df_1_complete.csv") |> 
  mutate(hum_cases = round(hum_cases))

county_comb_complete <- fread("df_cum_complete.csv") |> 
  mutate(hum_cases = round(hum_cases))

county_humanpop <- fread('county_humanpop.csv')
```

```{r Creating the proportion}

# Individual: Incidence
# The cattle and goat incidence were calculated 10M per population
df_indivi_inci <- county_indivi_complete |> 
  mutate(across(c(catt_incidence, goat_incidence), ~ round(.*10))) |> 
  mutate(human_incidence = round(human_incidence * 10)) |> 
  dplyr::select(date, county, contains("incidence")) |>
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .)))

# Individual: Proportion
df_indivi_inci_prop <- county_indivi_complete |> 
  mutate(across(c(catt_incidence, goat_incidence), ~ .*10)) |> 
  mutate(human_incidence = human_incidence/1e3) |> 
  dplyr::select(date, county, contains("incidence")) |>
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .))) |> 
  merge(county_humanpop, by = c("county", "date"))

# Individual: Cases
df_indivi_cases <- county_indivi_complete |> 
  dplyr::select(date, county, contains("cases")) |>
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .)))


# Combined : Incidence
df_comb_inci <- county_comb_complete |> 
  mutate(human_incidence = round(human_incidence*10),
         animal_incidence = round(animal_incidence*10)
         ) |> 
  dplyr::select(date, county, animal_incidence, human_incidence) |> 
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .)))


# combined: Proportion
df_comb_inci_prop <- county_comb_complete |> 
  mutate(human_incidence = human_incidence/1e3,
         animal_incidence = animal_incidence*10
         ) |> 
  dplyr::select(date, county, animal_incidence, human_incidence) |> 
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .)))|> 
  merge(county_humanpop, by = c("county", "date"))

# Combined: Cases
df_comb_cases <- county_comb_complete |> 
  dplyr::select(date, county, animal_cases, hum_cases) 
```

```{r Models for the mixed effects}

# Individual Incidence
indivi_models <- function(df, type, max_lag, ...) {
  # Type is either prop or incidence
  result_df <- tibble()
  if (type == "prop") {
    for (lag_value in 0:max_lag) {
      df_lagged <- df |>
        as_tibble() %>%
        mutate_at(
          vars(
            catt_incidence,
            goat_incidence
          ),
          list(~ lag(., n = lag_value))
        ) |>
        na.omit() |>
        mutate(date = as.Date(date))
      
      mod <-
        glmer(
          human_incidence ~ catt_incidence + goat_incidence + 
            (1 |
               county),
          data = df_lagged,
          family = binomial(link = "logit"),
          nAGQ = 0,
          weights = pop
        )
      
      mod_results <- broom.mixed::tidy(mod) %>%
        as_tibble() %>%
        mutate(
          term = case_when(
            term == "goat_incidence" ~ "Goat Incidence",
            term == "catt_incidence" ~ "Cattle incidence",
            term == "sd__(Intercept)" ~ "Sd - Random Intercept",
            TRUE ~ as.character(term)
          ),
          effect = case_when(
            effect == "ran_pars" ~ "Random",
            TRUE ~ effect
          ),
          variable = term
        ) |> 
        rename(`log odds` = estimate) |> 
        mutate(odds = exp(`log odds`),
               odds = ifelse(effect == "Random", NA, odds)
               ) |> 
        dplyr::select(1, 8,4, 9, 5:8) |> 
        group_by(variable) %>%
        mutate(
          conf_low = min(`log odds` - std.error * 1.645),
          conf_high = max(`log odds` + std.error * 1.645)
        ) %>%
        mutate(lag = lag_value)
      
      metrics <- broom.mixed::glance(mod) %>%
        dplyr::select(nobs, AIC, BIC)
      
      mod_results <- bind_cols(mod_results, metrics) |>
        mutate(across(
          c(
            `log odds`,
            odds,
            std.error,
            statistic,
            p.value,
            conf_low,
            conf_high,
            AIC,
            BIC
          ),
          ~ round(., 3)
        )) |>
        mutate(significance = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) 
      cat(paste("Runnning model for lag", lag_value), "\n")
      result_df <- bind_rows(result_df, mod_results)
    }
    
  } else if (type == "inci") {
    for (lag_value in 0:max_lag) {
      df_lagged <- df |>
        as_tibble() %>%
        mutate_at(
          vars(
            catt_incidence,
            goat_incidence
          ),
          list(~ lag(., n = lag_value))
        ) |>
        na.omit() |>
        mutate(date = as.Date(date))
      
      mod <-
         glmer.nb(
          human_incidence ~ catt_incidence + goat_incidence +
            (1 |
               county),
          data = df_lagged,
          nAGQ = 0
        )
      
      mod_results <- broom.mixed::tidy(mod) %>%
        as_tibble() %>%
        mutate(
          term = case_when(
            term == "goat_incidence" ~ "Goat Incidence",
            term == "catt_incidence" ~ "Cattle incidence",
            term == "sd__(Intercept)" ~ "Sd - Random Intercept",
            TRUE ~ as.character(term)
          ),
          effect = case_when(
            effect == "ran_pars" ~ "Random",
            TRUE ~ effect
          ),
          variable = term
        ) |> 
        rename(`log IRR` = estimate) |> 
        mutate(IRR = exp(`log IRR`),
               IRR = ifelse(effect == "Random", NA, IRR)
               ) |> 
        dplyr::select(1, 8,4, 9, 5:8) |> 
        group_by(variable) %>%
        mutate(
          conf_low = min(`log IRR` - std.error * 1.645),
          conf_high = max(`log IRR` + std.error * 1.645)
        ) %>%
        mutate(lag = lag_value)

      metrics <- broom.mixed::glance(mod) %>%
        dplyr::select(nobs, AIC, BIC)
      
      mod_results <- bind_cols(mod_results, metrics) |>
        mutate(across(
          c(
            `log IRR`,
            IRR,
            std.error,
            statistic,
            p.value,
            conf_low,
            conf_high,
            AIC,
            BIC
          ),
          ~ round(., 4)
        )) |>
        mutate(significance = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant"))
      cat(paste("Runnning model for lag", lag_value), "\n")
      result_df <- bind_rows(result_df, mod_results)
    }
    
  }
  
  return(result_df)

}

df_indivi_model_inci = indivi_models(df = df_indivi_inci, type = "inci", max_lag = 3)
df_indivi_model_prop = indivi_models(df = df_indivi_inci_prop, type = "prop", max_lag = 3)
write.csv(df_indivi_model_inci, "df_indivi_model_inci.csv")

# Combined Cases
comb_models <- function(df, type, max_lag, ...) {
  # Type is either prop or incidence
  result_df <- tibble()
  if (type == "prop") {
    for (lag_value in 0:max_lag) {
      df_lagged <- df |>
        as_tibble() %>%
        mutate_at(
          vars(
            animal_incidence,
          ),
          list(~ lag(., n = lag_value))
        ) |>
        na.omit() |>
        mutate(date = as.Date(date))
      
      mod <-
        glmer(
          human_incidence ~ animal_incidence + 
            (1 |
               county),
          data = df_lagged,
          family = binomial(link = "logit"),
          nAGQ = 0,
          weights = pop
        )
      
      mod_results <- broom.mixed::tidy(mod) %>%
        as_tibble() %>%
        mutate(
          term = case_when(
            term == "animal_incidence" ~ "Animal incidence",
            term == "sd__(Intercept)" ~ "Sd - Random Intercept",
            TRUE ~ as.character(term)
          ),
          effect = case_when(
            effect == "ran_pars" ~ "Random",
            TRUE ~ effect
          ),
          variable = term
        ) |> 
        rename(`log odds` = estimate) |> 
        mutate(odds = exp(`log odds`),
               odds = ifelse(effect == "Random", NA, odds)
               ) |> 
        dplyr::select(1, 8,4, 9, 5:8) |> 
        group_by(variable) %>%
        mutate(
          conf_low = min(`log odds` - std.error * 1.645),
          conf_high = max(`log odds` + std.error * 1.645)
        ) %>%
        mutate(lag = lag_value)
      
      metrics <- broom.mixed::glance(mod) %>%
        dplyr::select(nobs, AIC, BIC)
      
      mod_results <- bind_cols(mod_results, metrics) |>
        mutate(across(
          c(
            `log odds`,
            odds,
            std.error,
            statistic,
            p.value,
            conf_low,
            conf_high,
            AIC,
            BIC
          ),
          ~ round(., 3)
        )) |>
        mutate(significance = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant")) 
      cat(paste("Runnning model for lag", lag_value), "\n")
      result_df <- bind_rows(result_df, mod_results)
    }
    
  } else if (type == "inci") {
    for (lag_value in 0:max_lag) {
      df_lagged <- df |>
        as_tibble() %>%
        mutate_at(
          vars(
         animal_incidence
          ),
          list(~ lag(., n = lag_value))
        ) |>
        na.omit() |>
        mutate(date = as.Date(date))
      
      mod <-
         glmer.nb(
          human_incidence ~ animal_incidence +
            (1 |
               county),
          data = df_lagged,
          nAGQ = 0
        )
      
      mod_results <- broom.mixed::tidy(mod) %>%
        as_tibble() %>%
        mutate(
          term = case_when(
            term == "animal_incidence" ~ "Animal incidence",
            term == "sd__(Intercept)" ~ "Sd - Random Intercept",
            TRUE ~ as.character(term)
          ),
          effect = case_when(
            effect == "ran_pars" ~ "Random",
            TRUE ~ effect
          ),
          variable = term
        ) |> 
        rename(`log IRR` = estimate) |> 
        mutate(IRR = exp(`log IRR`),
               IRR = ifelse(effect == "Random", NA, IRR)
               ) |> 
        dplyr::select(1, 8,4, 9, 5:8) |> 
        group_by(variable) %>%
        mutate(
          conf_low = min(`log IRR` - std.error * 1.645),
          conf_high = max(`log IRR` + std.error * 1.645)
        ) %>%
        mutate(lag = lag_value)
      
      metrics <- broom.mixed::glance(mod) %>%
        dplyr::select(nobs, AIC, BIC)
      
      mod_results <- bind_cols(mod_results, metrics) |>
        mutate(across(
          c(
            `log IRR`,
            IRR,
            std.error,
            statistic,
            p.value,
            conf_low,
            conf_high,
            AIC,
            BIC
          ),
          ~ round(., 4)
        )) |>
        mutate(significance = ifelse(conf_low * conf_high > 0, "Significant", "Not Significant"))
      cat(paste("Runnning model for lag", lag_value), "\n")
      result_df <- bind_rows(result_df, mod_results)
    }
    
  }
  
  return(result_df)

}

df_combined_model_inci = comb_models(df = df_comb_inci, type = "inci", max_lag = 3)
df_combined_model_prop = comb_models(df = df_comb_inci_prop, type = "prop", max_lag = 3)
```

# Time Series Analysis

The following section contains the codes for time series analysis. Data was imported as follows;

```{r}
source('clean_tseries_data.R')
```

Then we test for stationarity as follows;\

```{r}
# Testing for stationary ---------------------------------------------------------------------------
# Test for human incidence
adf.test(df_complete1$human_incidence) # Not stationary

# Test for animal incidence
adf.test(df_complete1$animal_incidence)
```

\
Lag 3 had the highest correlation so we lagged the time series at lag 3.\
\

```{r}
# Lagging ------------------------------------------------------------

df2.1 <- df_complete1 |>
  mutate(date = as.Date(date),
         animal_incidence = lag(animal_incidence, 0)
         ) |>
  na.omit() |> 
  distinct(date, .keep_all = T)
adf.test(df2.1$human_incidence)


df_xts1 <- as.xts(x = cbind(df2.1$human_incidence, df2.1$animal_incidence), order.by = df2.1$date) |>
  setNames(c("human_incidence", "animal_incidence"))

plot(df_xts1)

# Data for train and testing -------------------------------------------------------------

df_xts <- window(df_xts1, start ='2014-04-01', end = '2022-12-01')
forecast.df <- window(df_xts1, start ='2023-01-01', end = '2023-12-01')

df2 <- df2.1 |> 
  filter(date < as.Date('2023-01-01'))
# Our training and testing data reaches 2022-12-01


```

## Testing for several aspects of the time series

### 1. Correlation of human incidence with it's past values

```{r}
# Correlation between human incidence and the previous lags
dd = df_xts |>
  data.frame(row.names = NULL) %>%
  mutate(date = zoo::index(df_xts) |>
           zoo::as.yearmon() |>
           yearmonth()) |>
  as_tsibble() 

dd |>
  gg_lag(human_incidence, geom = "point") +
  theme_light()
#> Lag 1 seems to have a strong correlation with the current value of human incidence

```

### 3. Decomposing the time series

```{r}
dcmp <- dd |>
  model(stl = STL(human_incidence))
```

Below is the trend and the time series

```{r}
# The trend
components(dcmp) |>
  as_tsibble() |>
  ggplot(aes(x = date)) +
  geom_xspline(aes(y = human_incidence, colour = "Human Incidence")) +
  geom_xspline(aes(y = trend, color = 'Trend')) +
  labs(y = "Incidence (human cases per 1000 pop)",
       title = "Human Brucellois incidence") +
  theme_light() +
  theme(
    axis.title = element_text(color = 'black'),
    axis.text = element_text(color = 'black'),
    plot.title = element_text(color = 'black', hjust = .5),
    legend.position = 'bottom'
  ) +
  scale_color_manual(values = c("Human Incidence" = "gray", "Trend" = "#D55E00")) +
  labs(color = 'Series:')
  
```

The time series seasonality

```{r}
# Seasonality
components(dcmp) |>
  as_tsibble() |>
  ggplot(aes(x = date)) +
  geom_xspline(aes(y = human_incidence, colour = "Human Incidence")) +
  geom_xspline(aes(y = season_year, color = 'Season-year')) +
  labs(y = "Incidence (human cases per 1000 pop)",
       title = "Human Brucellois incidence") +
  theme_light() +
  theme(
    axis.title = element_text(color = 'black'),
    axis.text = element_text(color = 'black'),
    plot.title = element_text(color = 'black', hjust = .5),
    legend.position = 'bottom'
  ) +
  scale_color_manual(values = c("Human Incidence" = "gray", "Season-year" = "#D55E00")) +
  labs(color = 'Series:')
```

All of them together

```{r}
# All of them together
components(dcmp) |>
  autoplot() +
  theme_light()
```

Plot without seasonality and trend

```{r}
# The plot without seasonality and trend
components(dcmp) |>
  as_tsibble() |>
  ggplot(aes(x = date)) +
  #geom_xspline(aes(y = human_incidence, colour = "Human Incidence")) +
  geom_xspline(aes(y = season_adjust - trend, color = 'Human Incidence without the seasonality')) +
  labs(y = "Incidence (human cases per 1000 pop)",
       title = "Human Brucellois incidence") +
  theme_light() +
  theme(
    axis.title = element_text(color = 'black'),
    axis.text = element_text(color = 'black'),
    plot.title = element_text(color = 'black', hjust = .5),
    legend.position = 'bottom'
  ) +
  scale_color_manual(values = c("Human Incidence" = "gray", "Human Incidence without the seasonality" = "#D55E00")) +
  labs(color = 'Series:')

```

## Train and testing data

We split our data into training and testing (80%, 20%) respectively as follows:\

```{r}
# Splitting ------------------------------------------------------------------------------

# train <- window(ts_diff, start = as.Date("2014-04-01"), end = "2021-12-01")
# test <- window(df_xts, start = as.Date("2022-01-01"), end = as.Date("2022-12-01"))

# We will take 80% for training, 20% for validation.
nrow(df_xts) # our data has 105 rows, thus 80% of this is 84 rows for training and 21 for testing.,
# 0% is upto 2021-03-01 and then the rest for testing
nrow_80 <- floor(.8 * nrow(df_xts))
train <- df_xts[1:nrow_80,]
test <- df_xts[(nrow_80 + 1):nrow(df_xts),]

```

The acf and pacf plot were as follows;\

```{r}
# ACF and PACF plots ---------------------------------------------------------------------

acf(train$human_incidence)

#> shows both trend and seasonality.
#> The slow decrease in the ACF as the lags increase is due to the trend, while the scalloped shape
#> is due to the seasonality. Time series that show no autocorrelation are called white noise. 

pacf(train$human_incidence)

```

## Training the Models

### 1. Model without the covariate

The model without covariate was trained as follows;\
\

```{r}
set.seed(123)
model_sarima <- train |>
  as.data.frame() %>%
  mutate(date = row.names(.) |>
           zoo::as.yearmon() |>
           yearmonth()) |>
  as_tsibble() |>
  model(
    ARIMA(human_incidence,
          ic = "aic",
          stepwise = T
          )
  ) |>
  report()
```

We then forecasted and extracted values for the forecasted as follows;

```{r}

original_data <- df_xts |>
  as.data.frame() %>%
  mutate(date = row.names(.) |> ymd())

fitted <- augment(model_sarima) |> 
  as.data.frame() %>%
  dplyr::select(date, fitted =.fitted, residuals = .resid) |> 
  mutate(date = ym(date) 
         ) 

CI <-  hilo(forecast::forecast(model_sarima, h = nrow(test))) |> 
  as.data.frame()

CI80 <- CI$`80%` |> 
  lapply(FUN = first) |> unlist() |>
  matrix(ncol = 3, byrow = T) |> 
  as.data.frame() |> 
  dplyr::select(-V3) |> 
  setNames(c("Lo80", "Hi80"))


CI95 <- CI$`95%` |> 
  lapply(FUN = first) |> unlist() |>
  matrix(ncol = 3, byrow = T) |> 
  as.data.frame() |> 
  dplyr::select(-V3) |> 
  setNames(c("Lo95", "Hi95"))

forecast_data_human_sarima <-
  data.frame(
    date = row.names(as.data.frame(test)) |>
      ymd(),
    PointForecast = CI$.mean,
    CI80,
    CI95
  )
```

The above data gave the following plot

```{r}
forecast_plot_human_sarima <- ggplot(original_data) +
  geom_ribbon(data = forecast_data_human_sarima, aes(date, ymin = Lo95, ymax = Hi95, fill = "95% CI"), alpha = 0.1) +
  geom_xspline(data = fitted, aes(date, y = fitted, color = "fitted")) +
  geom_ribbon(data = forecast_data_human_sarima, aes(date, ymin = Lo80, ymax = Hi80, fill = "80% CI"), alpha = .4) +
  geom_xspline(data = forecast_data_human_sarima, aes(date, PointForecast, colour = "Predicted human incidence")) +
  geom_xspline(aes(date, human_incidence, color = "Actual human incidence")) +
  theme_light() +
  theme(axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        legend.position = "bottom",
        plot.title = element_text(color = "black", hjust = .5)
        ) +
  xlab("Year") +
  ylab("Human Incidence") +
  scale_color_manual(
    values = c(
      "Predicted human incidence" = "red",
      "Actual human incidence" = "black",
      'fitted' = 'blue'
    )
  ) +
  scale_fill_manual(
    values = c(
      "95% CI" = "blue",
      "80% CI" = "blue"
    ),
        guide = guide_legend(
      override.aes = list(alpha = c(0.25, 1))
    )
  ) +
  labs(col = NULL, fill = NULL)
#plotly::ggplotly( forecast_plot_human_sarima)
forecast_plot_human_sarima

```

The model metrics were calculated as follows:\

```{r}

# Accuracy
test_acc <- test |>
  as.data.frame() %>%
  mutate(date = rownames(.) |> ymd() |> yearmonth()) |>
  as_tsibble() |>
  dplyr::select(-animal_incidence)
  
model_sarima_accuracy <- accuracy(forecast::forecast(model_sarima, h = nrow(test)), test_acc,
  measures = list(
    point_accuracy_measures,
    interval_accuracy_measures,
    distribution_accuracy_measures
  )
)

model_sarima_accuracy

```

### 2. Model with the covariate

The model without covariate was trained as follows;\

```{r}
set.seed(123)
model_sarima_covariates <- train |>
  as.data.frame() %>%
  mutate(date = row.names(.) |>
           zoo::as.yearmon() |>
           yearmonth()) |>
  as_tsibble() |>
  model(ARIMA(
    human_incidence ~ animal_incidence,
    ic = "aic",
    stepwise = T
  )) |>
  report()

```

We then forecasted and extracted values for the forecasted as follows;

```{r}
fitted_covariate <- augment(model_sarima_covariates) |> 
  as.data.frame() %>%
  dplyr::select(date, fitted =.fitted, residuals = .resid) |> 
  mutate(date = ym(date) 
  ) 

test_tsibble <- test |> 
  as.data.frame() %>%
  mutate(date = rownames(.) |> ymd() |> yearmonth()) |> 
  as_tsibble() |>
  dplyr::select(-human_incidence)
  

CI_covariate <- model_sarima_covariates |>
  forecast(new_data = test_tsibble) |> 
  hilo()

CI_covariate80 <- CI_covariate$`80%` |> 
  lapply(FUN = first) |> 
  unlist() |>
  matrix(ncol = 3, byrow = T) |> 
  as.data.frame() |> 
  dplyr::select(-V3) |> 
  setNames(c("Lo80", "Hi80"))


CI_covariate95 <- CI_covariate$`95%` |> 
  lapply(FUN = first) |> unlist() |>
  matrix(ncol = 3, byrow = T) |> 
  as.data.frame() |> 
  dplyr::select(-V3) |> 
  setNames(c("Lo95", "Hi95"))

forecast_data_covariate_sarima <-
  data.frame(
    date = row.names(as.data.frame(test)) |>
      ymd(),
    PointForecast = CI_covariate$.mean,
    CI_covariate80,
    CI_covariate95
  )

```

The above data gave the following plot

```{r}
forecast_plot_covariate_sarima <- ggplot(original_data) +
  geom_ribbon(data = forecast_data_covariate_sarima, aes(date, ymin = Lo95, ymax = Hi95, fill = "95% CI"), alpha = 0.1) +
  geom_ribbon(data = forecast_data_covariate_sarima, aes(date, ymin = Lo80, ymax = Hi80, fill = "80% CI"), alpha = .4) +
  geom_xspline(data = fitted_covariate, aes(date, y = fitted, color = "Fitted")) +
  geom_xspline(data = forecast_data_covariate_sarima, aes(date, PointForecast, colour = "Predicted human incidence")) +
  geom_xspline(aes(date, human_incidence, color = "Actual human incidence")) +
  theme_light() +
  theme(axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        legend.position = "bottom",
                plot.title = element_text(color = "black", hjust = .5)
        ) +
  xlab("Year") +
  ylab("Human Incidence") +
  scale_color_manual(
    values = c(
      "Predicted human incidence" = "red",
      "Actual human incidence" = "black",
      'Fitted' = 'blue'
    )
  ) +
  scale_fill_manual(
    values = c(
      "95% CI" = "blue",
      "80% CI" = "blue"
    ),
    guide = guide_legend(
      override.aes = list(alpha = c(0.25, 1))
    )
  ) +
  labs(col = NULL, fill = NULL)
forecast_plot_covariate_sarima
```

The model metrics were calculated as follows:

```{r}

model_sarima_covariates_coefficients <-
  tidy(model_sarima_covariates) |>
  dplyr::select(-.model) %>%
  as_tibble() %>%
  group_by(term) 


# Accuracy
test_acc2 <- test |>
  as.data.frame() %>%
  mutate(date = rownames(.) |> ymd() |> yearmonth()) |>
  as_tsibble() |>
  dplyr::select(-animal_incidence)

model_sarima_covariate_accuracy <-
  accuracy(
    model_sarima_covariates |>
      forecast(new_data = test_tsibble) ,
    test_acc,
    measures = list(
      point_accuracy_measures,
      interval_accuracy_measures,
      distribution_accuracy_measures
    )
  )

model_sarima_covariate_accuracy

```

### 3. Comparison of the Model with and Without a covariate

```{r}

model_sarima_accuracy2 <- model_sarima_accuracy |> 
    mutate(model = "Model without Covariate") |> 
  dplyr::select(-.model, -.type) |> 
  dplyr::select(model, RMSE, MAE, MAPE)

model_sarima_covariate_accuracy2 <- model_sarima_covariate_accuracy |> 
  mutate(model = "Model with Covariate") |> 
  dplyr::select(-.model, -.type) |> 
  dplyr::select(model, RMSE, MAE, MAPE)

metrics_df <- rbind(model_sarima_accuracy2, model_sarima_covariate_accuracy2)

kableExtra::kable(metrics_df)

write.csv(metrics_df, "metrics_df_lag0.csv", row.names = F)


#> We have saved the metrics df for several training set 70-30, 75-25, 80-20

```

The model with the covariate was better, based on the RMSE, MAE and MAPE. Thus, we used this to forecast the values for the year 2023,

## Full model for forecasting 2023 data

we fit a time series data with animal incidence as the covariate with data from 2014 to 2022 and then use it to forecast 2023 human brucellosis incidence. The model was fitted as follows;\
\

```{r}

# Time series so that we can be able to forecast 2023.
set.seed(123)
full.model <- df_xts |> 
  as.data.frame() %>%
  mutate(date = row.names(.) |>
           zoo::as.yearmon() |>
           yearmonth()) |>
  as_tsibble() |>
  model(ARIMA(
    human_incidence ~ animal_incidence,
    ic = "aic",
    stepwise = T
  )) |>
  report()

```

Then, the forecasted data is curated as follows;

```{r}
fitted_full <- augment(full.model) |> 
  as.data.frame() %>%
  dplyr::select(date, fitted =.fitted, residuals = .resid) |> 
  mutate(date = ym(date) 
  ) 

original_data2 <-   df_xts |>
  rbind(forecast.df) |> 
  as.data.frame() %>%
  mutate(date = row.names(.) |> ymd())

test_tsibble <- forecast.df |> 
  as.data.frame() %>%
  mutate(date = rownames(.) |> ymd() |> yearmonth()) |> 
  as_tsibble() |>
  dplyr::select(-human_incidence)


CI_full <- full.model |>
  forecast(new_data = test_tsibble) |> 
  hilo()

CI_full80 <- CI_full$`80%` |> 
  lapply(FUN = first) |> 
  unlist() |>
  matrix(ncol = 3, byrow = T) |> 
  as.data.frame() |> 
  dplyr::select(-V3) |> 
  setNames(c("Lo80", "Hi80"))


CI_full95 <- CI_full$`95%` |> 
  lapply(FUN = first) |> unlist() |>
  matrix(ncol = 3, byrow = T) |> 
  as.data.frame() |> 
  dplyr::select(-V3) |> 
  setNames(c("Lo95", "Hi95"))

forecast_data_full_sarima <-
  data.frame(
    date = row.names(as.data.frame(forecast.df)) |>
      ymd(),
    PointForecast = CI_full$.mean,
    CI_full80,
    CI_full95
  )
```

The above gave the following results\

```{r}

forecast_plot_full_sarima <- ggplot(original_data2) +
  geom_ribbon(data = forecast_data_full_sarima, aes(date, ymin = Lo95, ymax = Hi95, fill = "95% CI"), alpha = 0.1) +
  geom_ribbon(data = forecast_data_full_sarima, aes(date, ymin = Lo80, ymax = Hi80, fill = "80% CI"), alpha = .4) +
  geom_xspline(data = fitted_full, aes(date, y = fitted, color = "Fitted")) +
  geom_xspline(data = forecast_data_full_sarima, aes(date, PointForecast, colour = "Predicted human incidence")) +
  geom_xspline(aes(date, human_incidence, color = "Actual human incidence")) +
  theme_light() +
  theme(axis.text = element_text(color = "black"),
        axis.title = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        legend.position = "bottom",
        plot.title = element_text(color = "black", hjust = .5)
  ) +
  xlab("Year") +
  ylab("Human Incidence") +
  scale_color_manual(
    values = c(
      "Predicted human incidence" = "red",
      "Actual human incidence" = "black",
      'Fitted' = 'blue'
    )
  ) +
  scale_fill_manual(
    values = c(
      "95% CI" = "blue",
      "80% CI" = "blue"
    ),
    guide = guide_legend(
      override.aes = list(alpha = c(0.25, 1))
    )
  ) +
  labs(col = NULL, fill = NULL)
forecast_plot_full_sarima

```

The model coefficients are;

```{r}
full.model_coefficients <-
  tidy(full.model) |>
  dplyr::select(-.model) %>%
  as_tibble() %>%
  group_by(term)
knitr::kable(full.model_coefficients)
```

The results of the forecasted values versus the actual, together with the upper 95% CI were as follows;

```{r}
# Forecasted and Actual data
fore.actual <- forecast_data_full_sarima |> 
  mutate(Actual = as.vector(forecast.df$human_incidence),
         Date = as.Date(date) |> 
           zoo::as.yearmon() |>
           yearmonth()) |> 
  dplyr::select(Date = Date, 
         Forecasted = PointForecast,
         Actual,
         `Lower 95% CI` = Lo95,
         `Upper 95% CI` = Hi95,
         ) |> 
  mutate(across(where(is.numeric), ~as.numeric(round(., 3)))) 
knitr::kable(fore.actual)
```

## Supplementary Plots

```{r}

# Supplementary plots --------------------------------------------------------------------


forecast_plot_covariate_sarima_title <- forecast_plot_covariate_sarima +
  ggtitle("Forecast with exogenous variable")
forecast_plot_covariate_sarima_title
dev.off()
ggsave(
  "images/forecast_plot_covariate_sarima_title.png",
  width = 13,
  height = 6,
  dpi = 1e3
)

require(patchwork)
all <- forecast_plot_human_sarima/forecast_plot_covariate_sarima
all
dev.off()
ggsave(
  "images/all_without_title_sarima.png",
  width = 13,
  height = 10,
  dpi = 1e3
)

forecast_plot_human_sarima_title <- forecast_plot_human_sarima +
  ggtitle("Forecast without exogenous variable")
forecast_plot_human_sarima_title
all_title <- forecast_plot_human_sarima_title/forecast_plot_covariate_sarima_title
all_title
dev.off()
ggsave(
  "images/all_with_title_sarima.png",
  width = 13,
  height = 10,
  dpi = 1e3
)


```
