---
title: "Brucellosis Incidence Rate"
output:
  word_document: default
date: "2023-12-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
source("county_models.R")

```

## Descriptive Statistics

*The results reported in this file were coded in the* **county_models.R** *file*

Among Camels, the majority of cases (91.30%) are clinically confirmed, with a smaller proportion (8.70%) being lab confirmed; no cases were post-mortem. In the case of Cattle, clinical confirmation dominates (55.34%), followed by lab confirmation (41.26%), and a minor portion through post-mortem (3.40%). Goats show a significant reliance on clinical confirmation (82.16%), while lab confirmation accounts for 17.84%, and no post-mortem cases were reported. Humans exhibit a considerable reliance on lab confirmation (77.98%), with clinical confirmation comprising 22.02%, and no post-mortem cases. Sheep cases are predominantly clinically confirmed (79.27%), with lab-confirmed cases making up 20.73%, and no instances of post-mortem diagnosis. 

```{r}
table1.1
```

The descriptive statistics for the incidence rate of Brucellosis among Camels, Cattle, Goats, Humans, and Sheep are presented in the table 2. The human incidence was calculated per 1,000 population while the incidence for the other species was calculated per 1,000,000 population.

```{r table 2, fig.cap="Descritive Statistics for the Incidence Rate of Brucellosis among Camels, Cattle, Goats, Humans, and Sheep"}
table2
```

## Test for stationarity and Differencing

The Augmented Dickey-Fuller test was used to test for stationarity. The test was conducted for the data both at National and County level. At county level, the results were as shown below

```{r}
adf_res

```

At county level, all the variables were stationary at 5% level of significance. At national level, the results were as shown below

```{r}
adf_trend_res
```

At national level, the human incidence variable was not stationary at 5% level of significance. Thus, to make it stationary, the variables were differenced at lag 1

## Trend

The trend for all animal incidence rate combined was as shown below

```{r}
all_plus_hum
```


The trend for  humans as well as individual species was as shown below

```{r, fig.width=10, fig.height=10}
humans_plt/species_plt
```

## Spatial

The spatial distribution of the incidence rate of Brucellosis in animals combined was as shown below

\newpage

```{r, fig.width=20, fig.height=25}
animals
```

\newpage

The spatial distribution of the incidence rate of Brucellosis among Camels, Cattle, Goats, Humans, and Sheep are presented in the below,

\newpage

```{r, fig.width=20, fig.height=25}
all_plots
```

\newpage


## Correlation

The correlation between the incidence rate of Brucellosis among Camels, Cattle, Goats, Humans, and Sheep were calculated in different lags. There was a high correlation between cattle and goat incidence across all lags. This correlation plot was used to determine the best lag to test for association between the human incidence and animal incidences. The incidence rates for this plot were not differenced.

\newpage

```{r, fig.width= 22, fig.height=25}
all_cols
```

\newpage

## Models
### 1. Non Differenced Case
We first fit a time series linear model to the original data (original data implying non-differenced) at different lags. Lag 1 to Lag 6. The results of the models were as tabulated in the tables below. Note that the significance was calculated at **10% level of significance**.

In the folder,you can find the full data frames as follows;

1. *non_diff_individual.csv* - Results for non-differenced individual species model.
2. *non_diff_full.csv* - Results for non-differenced all species combined  species model.
3. *diff_individual.csv* - Results for differenced individual species model.
4. *diff_full.csv* - Results for differenced all species combined model.


```{r}
non_diff_indivi |>
  filter(lag == 0) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results for no lag (lag = 0)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |> 
  row_spec(0, bold = TRUE, background = "#1E90FF", color = "white") %>%
  kableExtra::kable_styling(full_width = FALSE)

non_diff_indivi |>
  filter(lag == 1) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    format = "pipe",
    caption = "Time Series Linear Model results at lag 1",
    pipe_options = "hold_position"
  ) |>
  kableExtra::kable_styling(full_width = FALSE) |>
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )


non_diff_indivi |>
  filter(lag == 2) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 2",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  kableExtra::kable_styling(full_width = FALSE) |>
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

non_diff_indivi |>
  filter(lag == 3) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 3",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

```

\newpage

```{r}
non_diff_indivi |>
  filter(lag == 4) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 4",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

non_diff_indivi |>
  filter(lag == 5) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 5",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

non_diff_indivi |>
  filter(lag == 6) |>
  select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 6",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

```

**A model was also fit for all the animal incidences combined, and at different lags (lag 1 to lag 6), as shown in the tables below;**

```{r}
non_diff_full |>
  filter(lag == 0) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results for no lag (lag = 0)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |> 
  row_spec(0, bold = TRUE, background = "#1E90FF", color = "white") %>%
  kableExtra::kable_styling(full_width = FALSE)

non_diff_full |>
  filter(lag == 1) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    format = "pipe",
    caption = "Time Series Linear Model results at lag 1",
    pipe_options = "hold_position"
  ) |>
  kableExtra::kable_styling(full_width = FALSE) |>
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )


non_diff_full |>
  filter(lag == 2) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 2",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  kableExtra::kable_styling(full_width = FALSE) |>
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

non_diff_full |>
  filter(lag == 3) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 3",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

```


```{r}
non_diff_full |>
  filter(lag == 4) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 4",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

non_diff_full |>
  filter(lag == 5) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 5",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

non_diff_full |>
  filter(lag == 6) |>
  select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 6",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

```

### 2. Differenced Case

Time series models were also fit for the differenced data, and at different lags (lag 1 to lag 6), as shown in the tables below;

```{r}
diff_indivi |>
  filter(lag == 0) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results for no lag (lag = 0) (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |> 
  row_spec(0, bold = TRUE, background = "#1E90FF", color = "white") %>%
  kableExtra::kable_styling(full_width = FALSE)

diff_indivi |>
  filter(lag == 1) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    format = "pipe",
    caption = "Time Series Linear Model results at lag 1 (differenced)",
    pipe_options = "hold_position"
  ) |>
  kableExtra::kable_styling(full_width = FALSE) |>
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )


diff_indivi |>
  filter(lag == 2) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 2 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  kableExtra::kable_styling(full_width = FALSE) |>
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

diff_indivi |>
  filter(lag == 3) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 3 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

```

\newpage

```{r}
diff_indivi |>
  filter(lag == 4) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 4 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

diff_indivi |>
  filter(lag == 5) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 5 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

diff_indivi |>
  filter(lag == 6) |>
  select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 6 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

```

**A model was also fit for all the animal incidences combined, differenced and at different lags (lag 1 to lag 6), as shown in the tables below;**

```{r}
diff_full |>
  filter(lag == 0) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results for no lag (lag = 0) (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |> 
  row_spec(0, bold = TRUE, background = "#1E90FF", color = "white") %>%
  kableExtra::kable_styling(full_width = FALSE)

diff_full |>
  filter(lag == 1) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    format = "pipe",
    caption = "Time Series Linear Model results at lag 1 (differenced)",
    pipe_options = "hold_position"
  ) |>
  kableExtra::kable_styling(full_width = FALSE) |>
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )


diff_full |>
  filter(lag == 2) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 2 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  kableExtra::kable_styling(full_width = FALSE) |>
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

diff_full |>
  filter(lag == 3) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 3 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

```


```{r}
diff_full |>
  filter(lag == 4) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 4 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

diff_full |>
  filter(lag == 5) |>
    select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 5 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

diff_full |>
  filter(lag == 6) |>
  select(-lag) |>
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 6 (differenced)",
    format = "pipe",
    pipe_options = "hold_position"
  ) |>
  #kableExtra::kable_styling(full_width = FALSE)
  row_spec(
    0,
    bold = TRUE,
    background = "#1E90FF",
    color = "white"
  )

```

## Model for Each specific County

### For individuL Species
Time series linear regression model was fit for all the counties giving the following results. Some counties didn't have results because they contained zero incidences in all the predictors (cattle, goat, sheep, and camel incidence). Also, the models have also been tested only at **lag 3**. The NA in the dataframe, indicates that the variable had zero incidence.

In the folder, you will find the results of the models as follows;

1. *individual_animal_incidence_per_county.csv* - Results for individual animal incidences per county (at lag 3).
2. *all_animal_incidence_per_county.csv* - Results for all animal incidences per county (at lag 3).

```{r, }
library(kableExtra)

coefficients_df %>%
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 3, for each county",
    format = "pipe"
    # centering = FALSE,
    # booktabs = TRUE,
    # longtable = TRUE,
  ) 
  # kableExtra::kable_styling(
  #   bootstrap_options = c("striped", "hover"),
  #   position = "left",
  #   full_width = F,
  #   fixed_thead = T,
  #   latex_options = c("striped", "HOLD_position", "repeat_header")
  # ) |>
  # row_spec(
  #   0,
  #   bold = TRUE,
  #   background = "#1E90FF",
  #   color = "white"
  # )
```


### For all the species combined

The model (at lag 3) for all the animal incidences combined were as follows;


```{r}
coefficients_df2 %>%
  knitr::kable(
    align = "l",
    caption = "Time Series Linear Model results at lag 3, for each county for all the animal incidences combined",
     format = "pipe"
    # centering = FALSE,
    # booktabs = TRUE,
    # longtable = TRUE,
  ) 
  # kableExtra::kable_styling(
  #   bootstrap_options = c("striped", "hover"),
  #   position = "left",
  #   full_width = F,
  #   fixed_thead = T,
  #   latex_options = c("striped", "HOLD_position", "repeat_header")
  # ) |>
  # row_spec(
  #   0,
  #   bold = TRUE,
  #   background = "#1E90FF",
  #   color = "white"
  # )
```


